---
title: "Estimating the impact of monetary policy shocks on different housing indicators: An SVar Approach"
author: "Yobin Timilsena"

execute:
  echo: false
  
bibliography: references.bib
---

> **Abstract.**  
>
> **Keywords.** bsvars, impulse responses, quarto, R, housing price index, monetary policy


# Introduction
Monetary policy is one of the key instruments used by central banks to influence the overall performance of an economy. In recent years, the housing market has become an increasingly important channel through which monetary policy affects the broader economy in Australia as much of individual savings are tied up in housing wealth. As such, understanding the relationship between monetary policy and the housing market is crucial for policymakers, investors, and households alike.

In this study, we aim to estimate the impact of monetary policy shocks on various housing indicators using a Structural Vector Autoregression (SVAR) approach. SVAR models are a popular econometric tool for investigating the dynamic interactions between macroeconomic variables. By applying this approach, we seek to shed light on how changes in monetary policy impact different dimensions of the housing market.

# Research Question


# Data and their Properties
## Empirical Project Setup
This project website is being developed as a **Quarto** document and the empirical work in conducted using `R`. The necessary datasets are imported from the Reserve Bank of Australia (RBA) and the Australian Bureau of Statistics (ABS) websites using `readrba` and `readabs` respectively.

```{r load packages, include=FALSE}
pkgs <- c('ggplot2', 'tidyr', 'dplyr', 'lubridate', 'readrba', 'readabs', 'tidyverse')
# install.packages(pkgs)
lapply(pkgs, library, character.only=TRUE)
```

The following is the data on **M1** money supply in the RBA website. 

```{r read m1 money supply, include=T}
m1 <-read_rba_seriesid("DMAM1N")
m1 %>%
  ggplot(aes(x = date, y = value)) +
  geom_line() +
  theme_minimal() +
  labs(title = "M1 (actual)")
```

# Model and Hypothesis




## References {.unnumbered}
