---
title: "Estimating the impact of monetary policy shocks on different housing indicators: A Bayesian SVAR Approach"
author: "Yobin Timilsena"

execute:
  echo: false
  
format: 
  html:
    toc-location: left
    
editor:
  markdown: 
    wrap: 72
  
bibliography: references.bib
---

> **Abstract.**  This project uses a Bayesian SVAR approach to estimate the effects of domestic and foreign monetary policy shocks on housing prices and the level of new housing construction in Australia. The identification relies on imposing exclusion-restrictions and the estimation process follows the @waggoner2003a algorithm using the Gibbs sampler. I build three extensions of the baseline model -- in the first two, I impose different types of hyperparameter prior distributions and estimate the hyperparameters of the priors of the model, while I incorporate common stochastic volatility in the third extension.  I find that a positive domestic monetary policy shock reduces both the number of new houses and housing prices, while a positive foreign (US) monetary policy shock reduces the number of new houses but increases housing prices.
>
> **Keywords.** bsvars, impulse responses, quarto, R, housing price index, monetary policy shocks, stochastic volatility, normal-gamma mixture prior, normal-inverse gamma prior.


# Introduction
Monetary policy is one of the key instruments used by central banks to influence overall economic activity. In recent years, the housing market has become an increasingly important channel through which monetary policy affects the broader economy (@AK_2017), particularly in Australia where around 57% of household wealth is tied up in the housing market (@Sweeney_2023). That number is thrice as big as the size of the super market in Australia and as such, understanding the relationship between monetary policy shocks and the housing market is crucial for policy-makers, investors, and households alike.

In this study, I aim to estimate the impact of monetary policy shocks on various housing indicators using a Structural Vector Autoregressive (SVAR) modelling approach with identification achieved through imposing exclusion restrictions. SVAR models are a popular econometric tool for investigating the dynamic interactions between macroeconomic variables. By applying this approach, I shed light on how changes in monetary policy impact different dimensions of the housing market. In particular, I find that a positive 25 basis point increase in the Australian Cash Rate reduces both the number of new houses and housing prices over the long run, while a positive foreign (US) monetary policy shock reduces the number of new houses but increases housing prices.

# Research Question
The objective of this paper is to estimate and quantify the impact of monetary policy shocks on different indicators of the housing market such as housing prices, housing transactions/sales, and home loan rates within an economy. Understanding these effects is crucial in understanding how money affects a key factor of the real side of the economy. I formulate this question within a SVAR framework and make use of @waggoner2003a algorithm to sample for posterior draws of the coefficient matrices. I further extend the model using common parameterisation of hierarchical priors and incorporating stochastic volatility, and estimate the model for better accuracy.

# Data and their Properties
## Empirical Project Setup
This project website is being developed as a **Quarto** document and the empirical work in conducted using `R`. The necessary datasets are imported from the Reserve Bank of Australia (RBA) and the Australian Bureau of Statistics (ABS) websites using `readrba` and `readabs` respectively.

```{r load packages, include=FALSE, warning=FALSE, eval=TRUE}
rm(list = ls())
pkgs <- c('ggplot2', 'tidyr', 'dplyr', 'lubridate', 'readrba', 'readabs', 'tidyverse', "xts", "plotly", "rmarkdown","quantmod", "zoo", "HDInterval", "plot3D", "mvtnorm")
# install.packages(pkgs)
lapply(pkgs, library, character.only=TRUE)
```


## Choice of variables
I use the following variables to answer this question. I discuss the relevance of each variable.

- $\log(M1)$: represents the log of the money supply M1. Both conventional and unconventional monetary policy shocks can change the stock of money supply and its size affects real variables of the economy.

- $\Delta CPI$: represents the year-on-year change in the Consumer Price Index (CPI). It is a measure of inflation in an economy and is affected by monetary policy shocks.

- $\log(c)$: represents the log of consumption of the economy. Monetary policy shocks can alter people's consumption-savings behaviour.

- $\log(GDP)$: represents the log of the Gross Domestic Product (GDP). Including this along with the consumption helps differentiate the effect on the non-consumption aspect of the economy.

- $loanrate$: represents the weighted average interest rates on owner-occupied home loans. This serves as a proxy for borrowing costs for households who save in the form of housing wealth.

- $AUCR$: represents the Australian Cash Rate Target. This is the major monetary policy instrument available to the RBA.

- $USFFR$: represents the Federal Funds Rate Maximum Target Rate. Monetary policy adopted in the US tend to ripple into other economies so this is a variable of interest. Another extension to this variable would be to include the Target rates of Australia's largest trading partners.

- $nhouses$: represents the number of new private dwellings (houses) approved for construction in Australia. Impact on housing prices might be dampened by the supply elasticity of housing captured by this variable. Moreover, macroeconomic conditions may determine the level of new construction projects undertaken domestically.

- $PPI$: represents the Property Price Index in Australia. The index is normalized with respect to the property prices in 2011-2012 (=100). Tracking the effect of monetary policy on housing prices is a key object of interest.


```{r download relevant variables, eval=FALSE, warning=FALSE}

# M1 Money Supply; quarterly; billions; log
m1.raw        <- read_rba_seriesid("DMAM1N")
m1.qtr        <- to.quarterly(xts(m1.raw$value, m1.raw$date), OHLC = F)
m1.qtr[,]     <- log(m1.qtr[,])

# Change in CPI; quarterly; pct change yoy
# Use cpi.raw <- read_rba_seriesid("GCPIAG") to get actual index value.
cpi.raw       <- read_rba_seriesid("GCPIAGYP")
cpi.qtr       <- to.quarterly(xts(cpi.raw$value, cpi.raw$date), OHLC = F)

# Consumption share of GDP; quarterly; millions -> billions; log
consumption.raw    <- read_rba_seriesid("GGDPECCVPSH")
consumption.qtr    <- to.quarterly(xts(consumption.raw$value, consumption.raw$date), OHLC = F)
consumption.qtr[,] <- log(consumption.qtr[,]/1000)

# Nominal GDP; quarterly; millions -> billions; log
gdp.raw       <- read_rba_seriesid("GGDPECCPGDP")
gdp.qtr       <- to.quarterly(xts(gdp.raw$value, gdp.raw$date), OHLC = F)
gdp.qtr[,]    <- log(gdp.qtr[,]/1000)

# Owner-occupied variable home loan rates; monthly; 
homeloan.raw  <- read_rba_seriesid("FILRHLBVS")
homeloan.qtr  <- to.quarterly(xts(homeloan.raw$value, homeloan.raw$date), OHLC = F)


# AUS CR Target Rate; monthly; starts from 1990
cashrate.raw  <- read_rba_seriesid("FOOIRATCR")
cashrate.qtr  <- to.quarterly(xts(cashrate.raw$value, cashrate.raw$date), OHLC = F)

# US FFR Max Target Rate; monthly -> quarterly,
usffr.raw     <- read_rba_seriesid("FOOIRUSFFTRMX")
usffr.qtr     <- to.quarterly(xts(usffr.raw$value, usffr.raw$date), OHLC = F)

# real money balances


# quantity of dwelling units, houses; monthly; jul 1983
nhomes.raw    <- read_abs_series("A418433F")
nhomes.qtr    <- to.quarterly(xts(nhomes.raw$value, nhomes.raw$date), OHLC = F)
nhomes.qtr[,] <- log(nhomes.qtr[,]/1000)

# Residential property price index; quarterly; sep 2003
homeprice.raw <- read_abs_series("A83728455L")
homeprice.qtr <- to.quarterly(xts(homeprice.raw$value, homeprice.raw$date), OHLC = F)
homeprice.qtr[,] <- log(homeprice.qtr[,])
```


```{r  truncating data; creating a dataframe, eval=FALSE, warning=FALSE}
df <- as.data.frame(na.omit(cbind(m1.qtr,cashrate.qtr,usffr.qtr)))
df                                      <- head(df,-1) # remove the last obs 
colnames(df)                            <- c("M1 Money Supply (logs)","RBA Target Cash Rate","US FFR Max Target Rate")
df["% Change in CPI, yoy"]              <- as.numeric(cpi.qtr[268:length(cpi.qtr)])                 
df["Aggregate Consumption (log)"]       <- as.numeric(tail(consumption.qtr,nrow(df)))
df["Nominal GDP (log)"]                 <- as.numeric(tail(gdp.qtr,nrow(df)))
df["Owner-occupied home loan rates"]    <- as.numeric(tail(homeloan.qtr,nrow(df)+1))[-133]
df["Quantity of new homes (log)"]                 <- as.numeric(tail(nhomes.qtr,nrow(df)+1))[-133]
df                                      <- head(df,-4) # drop the last 4 obs 
df["House Price Index (log)"]                 <- NA 
df[55:128,"House Price Index (log)"]          <- as.numeric(homeprice.qtr)
df                                      <- na.omit(df)
df                                      <- df %>% select("US FFR Max Target Rate", "Nominal GDP (log)", "Aggregate Consumption (log)",
                                                         "% Change in CPI, yoy", 
                                                         "RBA Target Cash Rate","M1 Money Supply (logs)", 
                                                         "Owner-occupied home loan rates",  "Quantity of new homes (log)",
                                                         "House Price Index (log)",)
                                            
```

```{r  Step 0 - Load saved data, eval=TRUE, include=FALSE, warning=FALSE}
  load(file = "ProjectData.RData") # load saved empirical data if needed
```

Below is a preview of the dataset used in this project.

```{r show the first few observations from the dataset, eval=TRUE, warning=FALSE}
head_df <- head(df)
rmarkdown::paged_table(head_df)
```

```{r save loaded data, eval=FALSE, warning=FALSE}
save.image(file="ProjectData.RData")
```


## Data Properties
The dataset consists quarterly data from 2003 Q3 to 2021 Q4. The variables discussed above are illustrated in the figure below. Note that the logged variables trend upwards because they are expressed in their levels, while variables expressed in percentage change terms do not exhibit this behaviour.

```{r plot the data, echo=FALSE}
#| label: fig-dataplot
#| fig-cap: "Summary of variables"
#| fig-cap-location: top

var_names <- colnames(df)
date_range <- as.yearqtr(rownames(df), format = "%Y Q%q")
par(mfrow = c(3,3), mar=c(2,2,2,2))
for (j in 1:ncol(df)){
    plot(x = date_range, y = df[,j],,type='l',
       main = paste(var_names[j]), ylab = "", xlab = "",
       lwd = 2.5,
       ylim = c(min(df[,j]),max(df[,j])))
}

```

```{r  save pdf of data plots, eval=FALSE}
pdf(file = "dataplot.pdf", width = 16, height = 9)
var_names <- colnames(df)
date_range <- as.yearqtr(rownames(df), format = "%Y Q%q")
par(mfrow = c(3,3), mar=c(2,2,2,2))
for (j in 1:ncol(df)){
  plot(x = date_range, y = df[,j],,type='l',
       main = paste(var_names[j]), ylab = "", xlab = "",
       lwd = 2.5,
       ylim = c(min(df[,j]),max(df[,j])))
}
dev.off()
```


### ADF Tests
I perform and display augmented Dickey-Fuller (ADF) test results on the variables. This test determines the order of integration of our variables. The null hypothesis of the test is that a selection set of time-series observations exhibit unit-root non-stationarity.

I report, for each variable, the difference level at which the ADF tests rejects the null that the series is non-stationary. It can be seen that most variables exhibit an order of integration 1. The variable for foreign monetary policy rate (the US Federal Funds rate) exhibits an order of integration 2 whereas house prices exhibits integration of order 0.

SVAR models generally incorporate an AR structure with multiple lags, so variables with lower orders of integration can be used as-is with these models.

```{r  perform adf tests on the data, warning= FALSE, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: ADF Tests"

perform_adf_tests <- function(df) {
  # Create an empty dataframe to store the results
  results <- data.frame(Variable = character(), TestType = character(), 
                        TestStatistic = numeric(), PValue = numeric(), 
                        stringsAsFactors = FALSE)

  # Iterate over each column in the dataframe
  for (col in colnames(df)) {
    # Remove NA values from the column
    column_data <- na.omit(df[[col]])

    # Perform ADF test for levels
    adf_levels <- tseries::adf.test(na.omit(column_data), k = 4)

    # Check if p-value is less than or equal to 0.05
    if (adf_levels$p.value <= 0.05) {
      results <- bind_rows(results,
        data.frame(Variable = col, TestType = "Levels", 
                   TestStatistic = adf_levels$statistic,
                   PValue = adf_levels$p.value)
      )
    } else {
      # Perform ADF test for first difference
      adf_diff1 <- tseries::adf.test(na.omit(diff(column_data)), k = 4)
      
      # Check if p-value is less than 0.05
      if (adf_diff1$p.value < 0.05) {
        results <- bind_rows(results,
          data.frame(Variable = col, TestType = "First Difference", 
                     TestStatistic = adf_diff1$statistic,
                     PValue = adf_diff1$p.value)
        )
      } else {
        # Perform ADF test for second difference
        adf_diff2 <- tseries::adf.test(na.omit(diff(column_data, differences = 2)), k = 4)
        
        results <- bind_rows(results,
            data.frame(Variable = col, TestType = "Second Difference", 
                       TestStatistic = adf_diff2$statistic,
                       PValue = adf_diff2$p.value)
          )
      }
    }
  }


  # Return the results dataframe
  return(results)
}

adf_test_results <- perform_adf_tests(df)
rmarkdown::paged_table(adf_test_results)

```
```{r save the adf test results as a latex table, eval=FALSE}
adf_table = xtable::xtable(adf_test_results)
print(adf_table, file = "adf_table.tex", include.rownames = FALSE)
```



### ACF/PACF Plots
Next, I display the Autocorrelation Function (ACF) and Partial Autocorrelation Function (PACF) plots for our selection of variables below.

The Autocorrelation Function (ACF) measures the correlation between a time series and its lagged values. It calculates the correlation coefficient between the time series and its lagged values at different lags. The ACF helps identify the presence of any significant autocorrelation in the data, indicating the degree of dependence between observations at different time points. ACF can help identify long-term trends and seasonality in the time series by examining the pattern of autocorrelation at different lags. 

The Partial Autocorrelation Function (PACF) measures the correlation between a time series and its lagged values, controlling for the intermediate lags. PACF is particularly useful for identifying the order of the autoregressive (AR) process in time series modeling, as it helps identify the lagged values that have a direct and significant influence on the current value.

```{r  ACF, include=TRUE}
#| label: ACF-Plots
#| fig-cap: "ACF Plots"
#| code-fold: true
#| code-summary: "Code: ACF Plots"
par(mfrow = c(3,3), mar=c(2,2,2,2))
for (j in 1:ncol(df)){
  acf(df[,j], xlab="")
  title(main=var_names[j])
}

```

```{r save the acf plots as pdf, eval=FALSE}
pdf(file = "acf.pdf", width = 12, height = 8)
par(mfrow = c(3,3), mar=c(2,2,2,2))
for (j in 1:ncol(df)){
  acf(df[,j], xlab="")
  title(main=var_names[j])
}
dev.off()
```

```{r  PACF, include=TRUE}
#| label: PACF-Plots
#| fig-cap-location: top
#| fig-cap: "PACF Plots"
#| code-fold: true
#| code-summary: "Code: PACF Plots"
par(mfrow = c(3,3), mar=c(2,2,2,2))
for (j in 1:ncol(df)){
  pacf(df[,j], xlab="")
  title(main=var_names[j])
}
```

```{r  save pacf plots in a pdf document, eval=FALSE}
pdf(file = "pacf.pdf", width = 12, height = 8)

par(mfrow = c(3,3), mar=c(2,2,2,2))
for (j in 1:ncol(df)){
  pacf(df[,j], xlab="")
  title(main=var_names[j])
}
```

The ACF plots for most variables show a gradually decreasing positive autocorrelation over time. The ACF for USFFR implies that policy decision reverse after about 10 quarters. The ACF for change in CPI and quantity of new both both indicate the presence of cycles in the data.

The PACF plots generally indicate that the first lag is really important in explaining the dynamics of the variables in our dataset. 

# Model and Hypothesis
I employ a structural VAR model to assess the impact of monetary policy shocks on different housing indicators. The SVAR model with $p$ lags can be written as
$$
\begin{align}
&B_0Y_t = B_0 + B_1 Y_{t-1} + \dots + B_p Y_{t-p} + U_t\\
&U_{t}| Y_{t-1} \dots Y_{t-p} \sim _{iid} ( 0_N, I_N)
\end{align}
$$ 
where
$$
Y_t =
            \begin{pmatrix}
                USFFR_t\\
                \log(GDP_t)\\
                \log(c_t)\\
                \Delta CPI_t\\
                AUCR_t\\
                \log(q_{\text{new homes}_t})\\
                \log(HPI_t)\\
                i_{\text{home loan}_t}\\
                \log(M1_t)
            \end{pmatrix}
$$

$B_0$ is a **structural matrix ** that captures the contemporaneous relationship between the variables in $Y_t$. $U_t$ represents conditionally independent **structural shocks**.

In many cases, the structural model can be estimated utilizing the information from its corresponding RF model
$$
\begin{align}
&Y_t = A_0 + A_1 Y_{t-1} + \dots + A_p Y_{t-p} + E_t\\
&E_{t}| Y_{t-1} \dots Y_{t-p} \sim _{iid} ( 0_N, \Sigma)
\end{align}
$$ 
where $A_i = B_0^{-1}B_i$ and $B_0^{-1}I_N  (B_0^{-1})'$.


The identification in the SVAR model can be achieved either by using some exclusion restrictions, sign restrictions, instrumental variables, or prior distribution. The next section will talk about the exact composition of the structural matrix and the conditions for identification. 

## Identification
I plan to use exclusion-restrictions to identify the structural matrix $B_0$ using the structural form equation. In particular, I will impose a lower-triangular restriction on $B_0$ and employ the solution concept in @waggoner2003b who use a normalization rule as an optimal solution to the local identification problem. I will then employ the Gibbs sampler for a SVAR model with exclusion restrictions as in @waggoner2003a to obtain draws for $B_0$ and $B_+$.

The structural form model, in matrix form, can be expressed as follows

$$ B_0 Y = B_+ X + U, \qquad \qquad U|X \sim \mathcal{MN}_{N \times T}(\textbf{0}_{N \times T}, I_T, I_N) $$
where 

$B_0$ is a $N \times N$ contemporaneous effects matrix.

$Y = [y_1, \dots, y_T]$ is a $N \times T$ matrix of observations.

$B_+ = [B_d, B_1, \dots, B_p]$ is a $N \times K$ matrix of autoregressive parameters, where $K = Np + d$  ($d$ is the number of deterministic terms;  $p$ is the number of lags).

$X = [x_1, \dots, x_T]$ is a $K \times T$ matrix of lagged observations where each $x_t = (1, y_{t-1}, \dots, y_{t-p})'$.

$U = [u_1, \dots, u_T]$ is a $N \times T$ matrix of structural shocks.

For convenience of coding and inference purposes, we consider a row-wise equation form as follows:
$$ B_{0[n.\cdot]} Y = B_{+n} X + U_n, \qquad \qquad U_n|X \sim \mathcal{N}(\mathbf{0}_T, I_T) $$

If $r_n$ denotes the number of elements in the $n^{th}$ row of $B_0$ that stay unrestricted, then we can further decompose $B_{0[n.\cdot]}$ into $b_n$ and $V_n$. 

$b_n$ is a $1 \times r_n$ vector of unrestricted elements in the $n^{th}$ row of $B_0$.

$V_n$ is a $r_n \times N$ matrix which places elements of $b_n$ in the appropriate positions to impose the restrictions on $B_0$. 

Then, the row-wise equation form can be written as follows:
$$ b_n V_n Y = B_{+n} X + U_n, \qquad \qquad U_n|X \sim \mathcal{N}(\mathbf{0}_T, I_T) $$
Following @waggoner2003a and @arias2018a, we define that $(B_+, B_0)$ follow jointly a Normal-Generalised Normal (NGN) distribution denoted as 
$$ p(B_+, B_0) \sim \mathcal{NGN}(B, \Omega, S, \nu)$$ 

if $B_{+n}$ follows a K-variate normal distribution given $b_n$ 
$$ p(B_{+n}|b_n) = \mathcal{N}_k(b_nV_nB, \Omega) $$  with kernel $$ p(B_{+n}|b_n) \propto \exp \left\{  -\frac{1}{2} \left( B_{+n} - b_nV_nB \right) \Omega^{-1}  \left( B_{+n} - b_nV_nB \right)' \right\} $$


for $n = 1, \dots, N$ and $b_1, \dots, b_N$ jointly have a distribution whose kernel is specified by 
$$ p(b_1, \dots, b_N) \propto | \det(B_0) |^{\nu - N} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N b_n V_n S^{-1} V_n' b_n'  \right\} $$

The joint-distribution of $(B_+, B_0)$ can then be written as
$$ p(B_0, B_+) = \left( \prod_{n=1}^N p(B_{+n}|b_n)\right) p(b_1, \dots, b_N) $$

This classification of the joint-distribution allows us to obtain natural-conjugate prior and corresponding posterior distributions.

Moreover, the following `ltexclusion` function imposes a lower-triangular restriction on $B_0$ and creates a list containing a matrix of $b_n$ and corresponding $V_n$ row vectors. 
```{r  Impose the type of exclusion restriction, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: Imposing lower-triangular exclusion restrictions"
ltexclusion = function(usedata){
  BM.V        = vector("list",usedata$N)
for (n in 1:usedata$N){
  BM.V[[n]]   = cbind(diag(n),matrix(0,n,usedata$N-n))
}

B0.initial    = matrix(0,usedata$N,usedata$N)
for (n in 1:usedata$N){
  unrestricted               = apply(BM.V[[n]],2,sum)==1
  B0.initial[n,unrestricted] = rnorm(sum(unrestricted))
}
B0Vlist       = list(B0.initial = B0.initial, V = BM.V)
}
```


## Baseline Model
### Prior distribution
Given this parameterisation, we can write down the kernel of the **prior** given hyperparameters $(\underline B, \underline \Omega, \underline S, \underline \nu)$ as follows:
$$ 
| \det(B_0) |^{\underline \nu - N} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N b_n V_n \underline S^{-1} V_n' b_n'  \right\}  \exp \left\{  -\frac{1}{2}  \sum_{n=1}^N  \left( B_{+n} - b_nV_n \underline B \right) \underline \Omega^{-1}  \left( B_{+n} - b_nV_n \underline B \right)' \right\}  
$$

This prior distribution has two key advantages. Firstly, it leads to a full-conditional posterior distributions that allow efficient sampling. This allows us to estimate the structural parameters of the SVAR model.

Secondly, it belongs to a class of reference prior distributions that are invariant to the pre-multiplication of the parameter matrices by a rotation matrix up to which the system is identified (see @rr2010). This allows us to conduct a proper Bayesian treatment of this model given the identification above.


```{r  include code from L16 codes.R}
  source("L16 codes.R")
```

#### Calibration of the prior
- $\underline \nu = N$ is a commonly chosen value as it implies that the generalised-normal part is equivalent to a $r_n$-variate normal with the mean equal to a vector of zeros and the covariance matrix equal to $\underline S$.
- $\underline S = \kappa_3 I_N$ implies that the covariances across the rows of $B_0$ is zero, and the variance of each row is homoskedastic (constant). $\kappa_3$ can be interpreted as a contemporaneous effects shrinkage and is set to 10.
- $\underline B = [0_{N \times 1} \; \kappa_4 I_N \; 0_{N \times (p-1)N}]$ implies an AR1 process for the structural VAR at the prior mean. In this calibration, $\kappa_4 = 1$, the AR1 process is a random walk process.
- $\underline \Omega = \begin{pmatrix} \kappa_2 & 0\\ 0 & \kappa_1 I_{Np} \end{pmatrix}$ is the prior covariance matrix. It is taken to be a diagonal matrix with the diagonal elements set as the Litterman prior. $\kappa_2$ represents the constant term shrinkage and is set to 10. $\kappa_1$ represents the autoregressive slope shrinkage and is set to 0.1.

We also calibrate the number of draws $S = 5000$ for any sampling, while the $S.burnin = 100$ represents the number of draws that are sampled first and then discarded.

The following `R` code creates a list of model parameters with the calibration as above. 
```{r  set prior parameters, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: List of Parameters"
# set the priors' parameters
parameters = list(
  kappa1    = .1,       # autoregressive slope shrinkage
  kappa2    = 10,       # constant term shrinkage
  kappa3    = 10,       # contemporaneous effects shrinkage
  kappa4    = 1,        # VAR prior persistence
  S         = 5000,     # number of sample draws
  S.burnin  = 100,      # number of initial draws that are burned-in
  h         = 16        # forecast horizon
)
```

The following `R` function takes as argument data and model parameters to compute parameters of the prior distribution and store it as a list.
```{r  define the prior function, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: Prior Function"
# A function that computes and stores 
# all the prior distribution components given a parameter list input
prior     = function(parameters, usedata){
  priors  = list(
  B       = cbind(rep(0,usedata$N), parameters$kappa4*diag(usedata$N), matrix(0, usedata$N, (usedata$p-1)*usedata$N)), # random walk prior
  Omega   = diag(c(parameters$kappa2,parameters$kappa1*((1:usedata$p)^(-2))%x%rep(1,usedata$N))),
  # Omega = diag(c(parameters$kappa2,parameters$kappa1*rep(1,usedata$N*usedata$p))),
  S       = parameters$kappa3*diag(usedata$N),
  nu      = usedata$N
  )
}
```



### Likelihood Function
The conditional normality of the error term allows us to write the kernel of the likelihood function and show that it can be expressed as a NGN distribution.
$$
\begin{align*}
    &L(B_+, B_0|Y,X) \propto | \det(B_0^{-1}B_0^{-1'})|^{-\frac{T}{2}} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N \left(b_n V_n Y - B_{+n} X \right) \left( b_n V_n Y - B_{+n} X  \right)' \right\}\\
    & = | \det(B_0)|^{T} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N \left( b_n V_n Y Y' V_n' b_n' - 2 b_n V_n Y X' B_{+n}' + B_{+n} X X' B_{+n}' \right) \right\}\\
    & = | \det(B_0)|^{T} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N \left( b_n V_n Y Y' V_n' b_n' + B_{+n} X X' B_{+n}' - 2 b_n V_n Y X' (XX')^{-1} (XX') B_{+n}' \right. \right. \\
    & \left. \left. + b_n V_n Y X' (XX')^{-1} (XX') (XX')^{-1} XY'V_n'b_n' - b_n V_n Y X' (XX')^{-1} (XX') (XX')^{-1} XY'V_n'b_n' \right) \right\}\\
    & = | \det(B_0)|^{T} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N \left( b_n V_n [ YY' - YX'(XX')^{-1}XY'] V_n' b_n'  + B_{+n} X X' B_{+n}' \right. \right. \\
    & \left. \left. - 2 b_n V_n Y X' (XX')^{-1} (XX') B_{+n}' + b_n V_n Y X' (XX')^{-1} XY' V_n' b_n' \right) \right\}\\
    & \text{}\\
    & \text{Let $\hat A = YX' (XX')^{-1} $, then we can simplify}\\
    & \text{}\\
    & =  | \det(B_0)|^{T} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N \left( b_n V_n [ YY' - \hat A XY'] V_n' b_n'  + B_{+n} X X' B_{+n}' - 2 b_n V_n \hat A (XX') B_{+n}' \right. \right.\\ & \left. \left. + b_n V_n \hat A XY' V_n' b_n' \right) \right\}\\
    & =  | \det(B_0)|^{T} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N \left( b_n V_n [ YY' - \hat A XY'] V_n' b_n'  + (B_{+n} - b_n V_n \hat A) X X' (B_{+n} - b_n V_n \hat A)' \right) \right\}\\
    & =  | \det(B_0)|^{T} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N b_n V_n [ YY' - \hat A XY'] V_n' b_n' \right\} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N (B_{+n} - b_n V_n \hat A) X X' (B_{+n} - b_n V_n \hat A)' \right\}
\end{align*}
$$

Comparing this with the general NGN kernel, we can write that

$$ 
L(B_+, B_0|Y, X) = \mathcal{NGN}(\tilde B, \tilde \Omega, \tilde S, \tilde \nu) 
$$
where

$$ 
\tilde B = \hat A, \quad \tilde \Omega = (XX')^{-1}, \quad \tilde S = ( YY' - \hat A XY')^{-1}, \quad \tilde \nu = T + N. 
$$
Hence, we have shown that the likelihood function follows a NGN distribution.

### Posterior Distribution
The prior and the likelihood can be used to obtain the posterior as follows:

$$
\begin{align*}
    & p(B_+, B_0|Y, X)  \propto L(B_+, B_0|Y, X) p(B_0, B_+)\\
    & = | \det(B_0)|^{T} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N \left(b_n V_n Y - B_{+n} X \right) \left( b_n V_n Y - B_{+n} X  \right)' \right\}\\
    & \times | \det(B_0) |^{\underline \nu - N} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N b_n V_n \underline S^{-1} V_n' b_n'  \right\}  \exp \left\{  -\frac{1}{2}  \sum_{n=1}^N  \left( B_{+n} - b_nV_n \underline B \right) \underline \Omega^{-1}  \left( B_{+n} - b_nV_n \underline B \right)' \right\} \\
    & =  | \det(B_0) |^{T + \underline \nu - N} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N \left(b_n V_n Y Y' V_n' b_n' - 2 b_n V_n Y X' B_{+n}' + B_{+n} X X' B_{+n}' \right) \right.\\
    & \left. + b_n V_n \underline S^{-1} V_n' b_n' + B_{+n} \underline \Omega^{-1} B_{+n}' - 2 b_n V_n \underline B \underline \Omega^{-1} B_{+n}' + b_n V_n \underline B \underline \Omega^{-1} \underline B' V_n' b_n' \right\}\\
    & =  | \det(B_0) |^{T + \underline \nu - N} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N b_n V_n [ YY' + \underline S^{-1} + \underline B \underline \Omega^{-1} \underline B'] V_n' b_n' - 2 b_n V_n [ YX' + \underline B \underline \Omega^{-1} ] B_{+n}' \right.\\
    & \left. + B_{+n} [ XX' + \underline \Omega^{-1} ] B_{+n}' \right\}\\ & \text{}\\
    & \text{ Let $ \bar \Omega = (XX' + \underline \Omega^{-1})^{-1} $, then we can write}\\ & \text{}\\
    & =  | \det(B_0) |^{T + \underline \nu - N} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N b_n V_n [ YY' + \underline S^{-1} + \underline B \underline \Omega^{-1} \underline B'] V_n' b_n' - 2 b_n V_n [ YX' + \underline B \underline \Omega^{-1} ] \bar \Omega \bar \Omega^{-1} B_{+n}' \right.\\
    & \left. + B_{+n} \bar \Omega^{-1} B_{+n}' \right\}\\
    & \text{}\\
    & \text{ Let $ \bar B = (YX' + \underline B \underline \Omega^{-1}) \bar \Omega $, then we can write}\\ & \text{}\\
    & =  | \det(B_0) |^{T + \underline \nu - N} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N b_n V_n [ YY' + \underline S^{-1} + \underline B \underline \Omega^{-1} \underline B' - \bar B \bar \Omega^{-1} \bar B' ] V_n' b_n'  \right.\\
    & \left.  + b_n V_n  \bar B \bar \Omega^{-1} \bar B' V_n' b_n' - 2 b_n V_n \bar B \bar \Omega^{-1} B_{+n}' + B_{+n} \bar \Omega^{-1} B_{+n}' \right\}\\
    & \text{}\\
    & \text{ Define $ \bar S = (YY' + \underline S^{-1} + \underline B \underline \Omega^{-1} \underline B' - \bar B \bar \Omega^{-1} \bar B')^{-1} $, then we can write}\\ & \text{}\\
    & =  | \det(B_0) |^{T + \underline \nu - N} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N b_n V_n \bar S^{-1} V_n' b_n'  + (B_{+n} - b_n V_n \bar B) \bar \Omega^{-1} (B_{+n} - b_n V_n \bar B)' \right\}\\
    &  = | \det(B_0) |^{T + \underline \nu - N} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N b_n V_n \bar S^{-1} V_n' b_n'  \right\} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N (B_{+n} - b_n V_n \bar B) \bar \Omega^{-1} (B_{+n} - b_n V_n \bar B)' \right\}
\end{align*}
$$
Thus, $$ p(B_+, B_0|Y,X) \sim \mathcal{NGN}(\bar B, \bar \Omega, \bar S, \bar \nu) $$
where the posterior parameters are as follows 
$$
\begin{align*}
&\bar B = (YX' + \underline B \underline \Omega^{-1}) \bar \Omega\\
& \bar \Omega = (XX' + \underline \Omega^{-1})^{-1} \\
& \bar S = (YY' + \underline S^{-1} + \underline B \underline \Omega^{-1} \underline B' - \bar B \bar \Omega^{-1} \bar B')^{-1}\\
&\bar \nu = T + \underline \nu
\end{align*}
$$

The following `R` function uses the prior and information contained in data to compute and store (as a list) the set of posterior parameters for our model. 
```{r  define the posterior function, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: Posterior Function"
posterior     = function(priors, usedata){
  Omega.inv   = solve(priors$Omega)
  Omega.post.inv = usedata$X%*%t(usedata$X) + Omega.inv
  Omega.post  = solve( Omega.post.inv )
  B.post      = (usedata$Y%*%t(usedata$X) + priors$B%*%Omega.inv) %*% Omega.post
  S.inv.post      = usedata$Y%*%t(usedata$Y) + solve(priors$S) + priors$B%*%Omega.inv%*%t(priors$B) - B.post%*%Omega.post.inv%*%t(B.post)
  nu.post     = ncol(usedata$Y) + priors$nu

  posteriors  = list(
  B           = B.post,
  Omega       = Omega.post,
  S.inv       = S.inv.post,
  nu          = nu.post
)
}

```

### (Gibbs) Sampling 
The sampling algorithm follows the derivation by @waggoner2003a. The algorithm first samples $b_1, \dots, b_n$ independent of $B_+$ in a serial, iterative fashion. Then, the sampled $b_1, \dots, b_n$ are normalised. Using the normalised $b_1, \dots, b_n$ and data, $B_{+n}$ can be drawn independently. 

#### Sample $b_1, \dots, b_n$ iteratively
Recall that the marginal posterior distribution of $b_1, \dots, b_n$ is proportional to
$$
| \det(B_0) |^{\underline \nu - N} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N b_n V_n \underline S^{-1} V_n' b_n'  \right\}
$$
The Gibbs sampler draws from the full conditional posterior distribution of vector $b_n$ given data as well as parameters from other rows of contemporaneous effects matrix, that is, $b_1, \dots,  b_{n-1}, b_{n+1}, \dots, b_{N}$. The full conditional posterior is denoted as
$$
p(b_n|Y, X, b_1, \dots,  b_{n-1}, b_{n+1}, \dots, b_{N})
$$
To sample from the full conditional posterior $p(b_n^{(s)}|Y, X, b_1^{(s)}, \dots,  b_{n-1}^{(s)}, b_{n+1}^{(s-1)}, \dots, b_{N}^{(s-1)})$ at each iteration $s$ and at each row $n$, the following steps are undertaken:

**Step 1:** define and compute the following values

- $U_n = chol \left( \bar \nu (V_n \bar S^{-1}V_n')^{-1}  \right)$ -- an $r_n \times r_n$ upper-triangular matrix;

- $w = B_{0[-n.\cdot]\perp}^{(s)}$ -- a $1 \times N$ matrix;

- $w_1 = wV_n'U-n' \cdot \left( w V_n'U_n'U_n V_n w' \right)^{-\frac{1}{2}}$ -- a $1 \times r_n$ vector;

- $W_n = [w_1' \quad w_{1 \perp}']'$ -- a $r_n \times r_n$ matrix.

**Step 2:** draw the elements of a $1 \times r_n$ vector $\alpha_n$ defined as follows:
- draw the first element by drawing $u \sim \mathcal{N}(\textbf{0}_{\nu+1}, \bar \nu^{-1}I_{\nu+1})$ and setting 
$$
\alpha_{n[\cdot.1]} = 
\begin{cases}
\sqrt{u'u} & \text{ with probability } 0.5\\
-\sqrt{u'u} & \text{ with probability } 0.5
\end{cases}
$$
- draw the remaining $r_n-1$ element of $\alpha_n$ from $\mathcal{N}(\textbf{0}_{r_n-1}, \bar \nu^{-1} I_{r_n-1})$.

**Step3:** compute the draw from the full-conditional posterior distribution of $b_n$ by 
$$
b_n^{(s)} = \alpha_n W_n U_n
$$

Note that $X\perp$ refers to the orthogonal-complement matrix of $X$. And, $B_{0[-n.\cdot]}$ refers to the matrix $B_{0}$ without its n$^{th}$ row.

#### Normalise $b_1, \dots, b_n$
@waggoner2003b provide a normalising rule that preserves the shape of the likelihood function. These normalised draws from the normal-generalised-normal posterior distribution are free of the local identification problem and hence the estimates post-normalisation are meaningful and appropriate for statistical inference.

**Step 1:** normalise with respect to one mode $\hat B_0$ of the posterior distribution.
Let $\hat B_0$ be defined as
$$
  \hat B_0 = chol((\bar\nu-N) \times \bar S)'
$$

**Step 2:** define scaling matrices $D_i$ for $i=1,\dots,2^N$.

These $N \times N$ matrices $D_i$ are diagonal matrices with diagonal elements equal to -1 or 1. Thus, $2^N$ set of $D_i$ matrices cover all possible combinations of -1 and 1 on the diagonal.

**Step 3:** compute the distance between $D_i B_0^{(s)}$ and $\hat B_0$.

$$
 d \left( \left[ \left( D_i B_0^{(s)} \right)^{-1'} - \hat B_0^{-1'}   \right] | (\hat B_0' \hat B_0)^{-1}   \right) = \sum_{n=1}^N \left[ \left( D_i B_0^{(s)} \right)^{-1'} - \hat B_0^{-1'}   \right]_{[n. \cdot]} (\hat B_0' \hat B_0)^{-1}  \left[ \left( D_i B_0^{(s)} \right)^{-1'} - \hat B_0^{-1'}   \right]_{[n. \cdot]}'
$$

The choice of $D_i$ that minimises this distance is used to create the normalised draw $D_{i*(s)}B_0^{(s)}$ after applying it to all of the $S$ draws.


#### Sample $B_{+n}$ independently
For each draw of $b_n^{(s)}$, a corresponding draw of $B_{+n}^{(s)}$ is directly sampled from the normal distribution below:
$$
B_{+n}^{(s)} \sim \mathcal{N}(b_n^{(s)}V_n \bar B_n, \bar \Omega)
$$


The `posteriorSimuations` function below takes model parameters, the posterior parameters, the exclusion restriction and the number of variables as given to sample draws using the Gibbs sampler. The functions `rgn`, `normalize.Gibbs.output.parallel` and `rnorm.ngn` are obtained from [Tomasz Woźniak](https://github.com/donotdespair)'s lecture notes. These functions are provided in the [Appendix](#appendix).
```{r  function to sample B0 and then B+ using wz2003, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: Sample B0 and Bplus Posteriors Function, Baseline Model"
posteriorSimulations  = function(parameters, posteriors, B0Vlist, N){
  t0                  = proc.time()
  B0.posterior        = rgn(n=parameters$S.burnin, S.inv=posteriors$S.inv, nu=posteriors$nu, V=B0Vlist$V, B0.initial=B0Vlist$B0.initial)
  t1                  = proc.time()
  (t1-t0)/60
  
  # sampling B0 from the posterior distribution using Gibbs
  t0                  = proc.time()
  B0.posterior        = rgn(n=parameters$S, S.inv=posteriors$S.inv, nu=posteriors$nu, V=B0Vlist$V, B0.initial=B0.posterior[,,parameters$S.burnin])
  t1                  = proc.time()
  (t1-t0)/60
  
  # normalisation
  B0.hat              = t(chol((posteriors$nu-N)*posteriors$S))                   # normalisation using this B0.hat should work
  BM.B0.posterior     = normalize.Gibbs.output.parallel(B0.posterior,B0.hat=B0.hat)
  t2                  = proc.time()
  (t2-t1)/60
  
  # sample B+ from the normal conditional posterior
  t2                  = proc.time()
  BM.Bp.posterior     = rnorm.ngn(BM.B0.posterior, B=posteriors$B,Omega=posteriors$Omega)
  t3                  = proc.time()
  (t3-t2)/60
  
  list(B0.posterior   = BM.B0.posterior, Bp.posterior = BM.Bp.posterior)
}
```



The `posteriorMeans` function below computes the sample average of the posterior draws of $B_0$ and $B_+$ matrices obtained from each sampling step.
```{r  function to compute means for posterior B0 and B+, echo=TRUE}
#| code-fold: true
#| code-summary: "posteriorMeans Function"
posteriorMeans = function(Bposteriors){
  Bposteriors.means = list(
  B0 = rowMeans(Bposteriors$B0.posterior, dims = 2),
  Bp = rowMeans(Bposteriors$Bp.posterior, dims = 2)
  )
}
```

The `pmatrix` function below prints the matrices in `R` in LaTeX form.
```{r  function to print matrix in latex format, echo=TRUE}
#| code-fold: true
#| code-summary: "Function to print matrix in LaTeX format"
pmatrix <- function(x) {
  if (is.matrix(x)) {
    cat(c("$$\\begin{equation*}\n",
          "\\left(",
          knitr::kable(x, format = "latex",
                       tabular = "array",
                       vline = "",
                       align = "c",
                       linesep = "",
                       toprule = NULL,
                       bottomrule = NULL),
          "\\right)\\, .\n",
          "\\end{equation*}$$\n"))
  } else {
    cat(c("$$\\begin{equation*}\n",
          x,
          "\\end{equation*}$$\n"))
  }
}

```

### Simulation Run
The purpose of this simulation run is to verify whether the model and corresponding code can replicate the true parameters of a data-generating process. To do this, I create artificial data containing 1000 observations simulated from a bi-variate Gaussian random walk process with the covariance matrix equal to the identity matrix of order 2. Then, I estimate a model with a constant term and 1 lag with the artificial data. And I show that the posterior mean of the autoregressive and the covariance matrices are close to an identity matrix and that the posterior mean of the constant term is close to a vector of zeros.

This process is laid out in steps below:

***Step 1:*** The following code generates artificial data containing 1000 observations from a bi-variate Gaussian random walk process with the covariance matrix equal to the identity matrix of order 2.
```{r  Step 1 - Generate artificial data, echo=TRUE}
#| echo: true
#| code-fold: true
#| code-summary: "Code: Generate artificial data"
set.seed(12345)

sim.p = 1
sim.T = 1000
sim.N = 2
sim.K = 1 + sim.N*sim.p

sim.Y           = arima.sim(list(order = c(0,1,0)), n = sim.T + sim.p-1, mean = 0, sd =1)
for (i in 2:sim.N){
  sim.Y         = rbind(sim.Y, arima.sim(list(order = c(0,1,0)), n = sim.T + sim.p-1, mean = 0, sd = 1))
}

sim.X           = matrix(1,1,sim.T)
for (i in 1:sim.p){
  sim.X         = rbind(sim.X, sim.Y[,(sim.p+1-i):(ncol(sim.Y)-i)])
}
sim.Y           = sim.Y[,-sim.p]
artificialdata  = list(p = sim.p, N = sim.N, K = sim.K, Y = sim.Y, X = sim.X)

```



***Step 2:*** We obtain a list of simulation priors and posteriors using the `prior` and `posterior` functions. 
```{r  Step 2 - Compute priors and posteriors for simulated data, echo=TRUE}
sim.priors      = prior(parameters, artificialdata)
sim.posteriors  = posterior(sim.priors, artificialdata)
```

***Step 3:*** We create a list of $V_n$ and $b_n$ corresponding to a lower triangular exclusion restriction on $B_0$ using the `ltexclusion` function.
```{r  Step 3 - Impose exclusion restriction on simulated data; get b_n and V_n, echo=TRUE}
sim.B0Vlist   = ltexclusion(artificialdata)
```

***Step 4:*** We sample the $B_0$ and $B_+$ posteriors with the Gibbs sampler using the `posteriorSimulations` function, and save the results for future use.
```{r  Step 4 - estimate and draw B0 and B+, echo=TRUE, eval=FALSE }
sim.Bposteriors       = posteriorSimulations(parameters, sim.posteriors, sim.B0Vlist, artificialdata$N)
```

```{r  save  simulated data}
#| eval: false
save(sim.Bposteriors, sim.priors, sim.posteriors, file = "sim-posteriors.RData")
```

```{r  Step 4b - load saved simulated posteriors, eval=TRUE}
load(file = "sim-posteriors.RData")
```

***Step 5:*** We compute the sample averages of our posterior $B_0$ and $B_+$. We use the `pmatrix` function to display the results as a matrix.
```{r  Step 5a - Compute the posterior means of B0 for simulated data, echo=TRUE, results='asis'}
sim.Bposteriors.means = posteriorMeans(sim.Bposteriors)
sim_B0 = pmatrix(sim.Bposteriors.means$B0)
```
We can see that the computed $B_0$ covariance matrix is numerically identical to an identity matrix.

```{r  Step 5b - Compute the posterior means of Bp for simulated data, echo=TRUE, results='asis'}
sim_Bp = pmatrix(sim.Bposteriors.means$Bp)
```
The first column of $B_+$ represents the posterior mean of the constant term. The values are small and close to zero.
The rest of the $B_+$ matrix represents the autoregressive matrix. Its posterior mean is numerically equal to an identity matrix.


### Data Results

***Step 1:*** I set the desired number of lags in the model and use the data to create matrices $X$ and $Y$. I store my data as a list named `mydata`.
```{r  Step 1 - set lags, create Y and X matrices, echo=TRUE}
#| echo: true
#| code-fold: true
#| code-summary: "Code: List of Data Matrices"

# Y is N by T; X is K by T

p = 4                                   # set a number of lags included
N = ncol(df)
K = 1 + N*p

Y = t(df[(p+1):nrow(df),])
X = matrix(1,1,ncol(Y))

for (i in 1:p){
  X    = rbind(X,t(df[((p+1):nrow(df))-i,]))
}

mydata = list(p=p,N=N,K=K,Y=Y,X=X)
```

***Step 2:*** We obtain a list of data priors and posteriors using the `prior` and `posterior` functions.
```{r  Step 2 - Compute priors and posteriors, echo=TRUE}
priors     = prior(parameters, mydata)
posteriors = posterior(priors, mydata)
```

***Step 3:*** We create a list of $V_n$ and $b_n$ corresponding to a lower triangular exclusion restriction on $B_0$ using the `ltexclusion` function.
```{r  Step 3 - Impose exclusion restrictions, echo=TRUE}
B0Vlist       = ltexclusion(mydata)
```


***Step 4:*** We sample the $B_0$ and $B_p$ posteriors with the Gibbs sampler using the `posteriorSimulations` function, and save the results for future use.
```{r  Step 4 - Estimate and sample B0 and B+, echo=TRUE,eval=FALSE}
Bposteriors           = posteriorSimulations(parameters, posteriors, B0Vlist, mydata$N)
```

```{r save posterior data, eval=FALSE}
save(Bposteriors, priors, posteriors, file = "data-posteriors.RData")
```


```{r  Step 4b - load saved data posteriors, eval=TRUE}
load(file = "data-posteriors.RData")
```

***Step 5:*** We compute the sample averages of our posterior $B_0$ and $B_+$. We use the `pmatrix` function to display the results as a matrix.
```{r  Step 5 - Compute posterior means for B0 and B+, echo=TRUE, results='asis'}
Bposterior.means = posteriorMeans(Bposteriors)
data_B0 = pmatrix(Bposterior.means$B0)
```

```{r  Step 5b - Compute the posterior means of Bp for actual data, echo=TRUE, results='asis'}
data_Bp = pmatrix(Bposterior.means$Bp)
```



# Extended Model
In the extended model, I will estimate the hyperparameters rather than setting them exogenously. Such estimation procedure often improves the fit of the model especially because the results can be sensitive to the parameterisation of the hyperparameters. 
In particular, I estimate $(\kappa_0, \kappa_+)$ such that $\underline S = \kappa_0 I_N$ and $\underline \Omega = \kappa_+ I_K$. I assume 

## Analysis of priors
In my extensions, I will assume the priors of hyperparameters $(\kappa_0, \kappa_+)$ to follow either an Inverse-Gamma 2 $(\mathcal{IG}2)$ or a Gamma $(\mathcal{G})$ distribution. In this section, I want to illustrate the properties of these distributions. Suppose that
$$
\begin{align*}
    \kappa_0 | \underline s_{\kappa_0}, \underline \nu_{\kappa_0} \sim \mathcal{IG}2(\underline s_{\kappa_0}, \underline \nu_{\kappa_0})  && \kappa_1 | \underline s_{\kappa_1}, \underline \nu_{\kappa_1} \sim \mathcal{G}(2 \underline s_{\kappa_1}, \frac{1}{2} \underline \nu_{\kappa_1}) 
  \end{align*}
$$
```{r analysis of priors}
#| code-fold: true
#| code-summary: "Code: Analysis of Priors"
#| include: true
#| echo: true
kappa0.s = .1
kappa0.nu = 1
kappap.s = .1/2
kappap.nu = 1 * 2
nDraws = 5000
  
kappa0.priorDraws = kappa0.s / rchisq(nDraws, df = kappa0.nu)
kappap.priorDraws = rgamma(nDraws, shape = kappap.s, scale = kappap.nu)

par(mfrow=c(1,2))
hist(kappa0.priorDraws,
     breaks = 100,
     col = "deepskyblue2",
     border = 'black',
     main = expression(paste("Distribution of ", kappa[0], " draws")),
     xlab = ""
     )
hist(kappap.priorDraws,
     breaks = 100,
     col = 'deepskyblue2',
     border = 'black',
     main = expression(paste("Distribution of ", kappa["+"], " draws")),
     xlab = ""
     )

```
We can see that the draws from the $\mathcal{IG}2$ are more positively skewed, have greater dispersion but thinner tails compared to draws from the $\mathcal{G}$ distribution. The dispersion is orders of magnitude higher for the $\mathcal{IG}2$ draws. 

## Extension 1: Normal-Gamma Prior Hyperparameter Estimation

### Prior Distribution
I postulate the following prior distributions for the hyperparameters:
$$
\begin{align*}
    \kappa_0 | \underline s_{\kappa_0}, \underline \nu_{\kappa_0} \sim \mathcal{IG}2(\underline s_{\kappa_0}, \underline \nu_{\kappa_0})  && \kappa_+ | \underline s_{\kappa_+}, \underline \nu_{\kappa_+} \sim \mathcal{G}(2 \underline s_{\kappa_+}, \frac{1}{2} \underline \nu_{\kappa_+}) 
  \end{align*}
$$

The `ext.priors` function creates a list containing all the elements of the prior.
```{r  hyperparameter priors, echo=TRUE}
#| code-fold: true
#| code-summary: "Extension Model Priors Function"
ext.priors = function(parameters,usedata){
  ext.prior = list(
  kappa0.s = .1,
  kappa0.nu = 1,
  kappap.s = .1,
  kappap.nu = 1,
  B       = cbind(rep(0,usedata$N), diag(usedata$N), matrix(0, usedata$N, (usedata$p-1)*usedata$N)), # random walk prior
  Omega   = parameters$kappa2 * diag(usedata$K),
  S       = parameters$kappa3*diag(usedata$N),
  nu      = usedata$N
)
}
```


Moreover, we have 
$$
  \begin{align*}
    p(B_{+n}|b_n, \kappa_+) = \mathcal{N}_k (b_n V_n B, \kappa_+ \Omega) && p(b_n| \kappa_0) = \mathcal{N}_{r_n}(0, \kappa_0 S)
  \end{align*}
$$

### Posterior Distribution

Thus, posteriors can be written as
$$
  \begin{align*}
    & p(\kappa_0 | Y, X, B_0) \propto p(B_0|\kappa_0) p(\kappa_0 | \underline s_{\kappa_0}, \underline \nu_{\kappa_0} )\\
    & \propto \prod_{n=1}^{N}\left(\kappa_0^{-\frac{1}{2}}\right)^{r_n} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N b_n V_n (\kappa_0 I_N)^{-1} V_n' b_n'  \right\} \cdot \kappa_0^{-\frac{\underline \nu_{\kappa_0}+2}{2}} \exp \left\{  -\frac{\underline s_{\kappa_0}}{2 \kappa_0} \right\}\\ 
    & \propto \kappa_0^{ -\frac{\underline \nu_{\kappa_0} + 2 + \sum_{n=1}^{N}r_n}{2}} \cdot \exp \left\{  -\frac{1}{2 \kappa_0} \sum_{n=1}^N b_n V_n V_n' b_n' + \underline s_{\kappa_0} \right\}\\
  \end{align*}
$$
  This gives
$$\bar s_{\kappa_0} =  \sum_{n=1}^N b_n V_n V_n' b_n' + \underline s_{\kappa_0}.$$
$$\bar \nu_{\kappa_0} = \underline \nu_{\kappa_0} +\sum_{n=1}^{N}r_n$$
  Also,
$$
  \begin{align*}
    & p(\kappa_+ | Y, X, B_+, B_0) \propto p(B_+|\kappa_+, B_0) p(\kappa_+ | \underline s_{\kappa_+}, \underline \nu_{\kappa_+} )\\
    & \propto \kappa_+^{-\frac{KN}{2}} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N (B_{+n} - b_n V_n \underline B) (\kappa_+ I_{K})^{-1} (B_{+n} - b_n V_n \underline B)'  \right\} \cdot \kappa_+^{\frac{\underline \nu_{\kappa_+} - 2}{2}} \exp \left\{  -\frac{ \kappa_+}{2 \underline s_{\kappa_+}} \right\}\\
    & = \kappa_+^{-\frac{-\underline \nu_{\kappa_+} + KN}{2} - 1} \cdot \exp \left\{  -\frac{1}{2} \left( (B_{+n} - b_n V_n \underline B)  (B_{+n} - b_n V_n \underline B)' \cdot \frac{1}{\kappa_+} + \frac{1}{\underline s_{\kappa_+}} \kappa_+  \right) \right\}\\
  \end{align*}
$$
  This gives
$$\bar\lambda = -\frac{-\underline \nu_{\kappa_+} + KN}{2}$$
$$\bar \chi =  \sum_{n=1}^N  (B_{+n} - b_n V_n \underline B)  (B_{+n} - b_n V_n \underline B)'$$
$$\bar \Psi = \frac{1}{\underline s_{\kappa_+}}$$
  
  
The `init.struct` function creates and initialise a list that contains matrices to store draws of $\{\kappa_0^{(s)}, \kappa_+^{(s)}, B_0^{(s)}, B_+^{(s)}\}_{s =1}^{S}$.
```{r  Structures, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: Create structure to store results"
  init.struct = function(usedata,S){
    kappa0 = array(NA,c(1,S))
    kappa0[1] = 10
    kappap = rep(NA, S)
    kappap[1] = 10
    B0.posterior = array(NA, c(usedata$N,usedata$N,S))
    Bp.posterior = array(NA, c(usedata$N,usedata$K,S))
    list(kappa0 = kappa0, kappap = kappap, B0.posterior = B0.posterior, Bp.posterior = Bp.posterior)
  }
```

### (Gibbs) Sampling
Given the full conditional posteriors of both $(\kappa_0,\kappa_+)$ and $(B_0, B_+)$, we can implement the Gibbs sampler as follows:

- Step 1: Initialise $(\kappa_0^{(0)},\kappa_+^{(0)})$. 

- Step 2: Use $(\kappa_0^{(s-1)},\kappa_+^{(s-1)})$ values to compute posterior parameters for each draw $(\bar B, \bar \Omega, \bar s, \bar \nu)^{(s)}$.

- Step 3: draw $(B_0, B_+)^{(s)} \sim \mathcal{NGN}(\bar B, \bar \Omega, \bar s, \bar \nu)^{(s)}$ using the sampling procedure presented in the baseline model.

- Step 4: draw $\kappa_0^{(s)} \sim p(\kappa_0 | Y, X, B_0^{(s)}) = \mathcal{IG2}(\bar s_{\kappa_0}, \bar \nu_{\kappa_0})$ and $\kappa_+^{(s)} \sim p(\kappa_+ | Y, X, B_0^{(s)}, B_+^{(s)}) = \mathcal{GIG}(\bar \lambda, \bar \chi, \bar \Psi)$.

Repeat steps 2-4 to get desired sample size and burn in some initial sample observations as needed.

The sampling function `ext.sampling` is given below. 
```{r  Sampling, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: Sampler for Extension 1"
ext.sampling = function(parameters, struct, priors, usedata){
  set.seed(12345)
  B0Vlist.initial = ltexclusion(usedata)
  B0.initial = B0Vlist.initial$B0
  for (i in 1:(parameters$S + parameters$S.burnin)){
    
    # Computing posterior parameters for each draw
    Omega.inv      = solve(struct$kappap[i] * priors$Omega)
    Omega.post.inv = usedata$X%*%t(usedata$X) + Omega.inv
    Omega.post     = solve(Omega.post.inv)
    B.post         = (usedata$Y%*%t(usedata$X) + priors$B%*%Omega.inv) %*% Omega.post
    S.post         = usedata$Y%*%t(usedata$Y) + solve(struct$kappa0[i] * priors$S) + priors$B%*%Omega.inv%*%t(priors$B) - B.post%*%Omega.post.inv%*%t(B.post) 
    nu.post        = ncol(usedata$Y) + priors$nu


  if (i > 1){
    B0.initial = struct$B0.post[,,i-1]
  }
  
  B0.i = rgn(n=1, S.inv = S.post, nu = nu.post, V = B0Vlist.initial$V, B0.initial = B0.initial)
  # B0.hat = t(chol((nu.post - usedata$N)*solve(S.post)))
  # B0.norm.i = normalize.Gibbs.output.parallel(B0.i, B0.hat)
  Bp.i = rnorm.ngn(B0.i, B.post, Omega.post)
  
  struct$B0.posterior[,,i] = B0.i
  struct$Bp.posterior[,,i] = Bp.i
  
  # Now, update kappa
  kappa0.nu.post = priors$kappa0.nu + (usedata$N/2)
  kappap.lambda.post = -(-priors$kappap.nu + usedata$K * usedata$N)/2
  kappap.psi.post = 1/priors$kappap.s
  kappa0.s.post = priors$kappa0.s
  for (n in 1:usedata$N){
    kappa0.s.post = kappa0.s.post + sum(B0.i[n,,1]^2)
    kappap.chi.post = (Bp.i[n,,1] - B0.i[n,,1] %*% B.post) %*% t(Bp.i[n,,1] - B0.i[n,,1] %*% B.post)
  }
  
  if (i != (parameters$S + parameters$S.burnin)){
    struct$kappa0[i+1] = kappa0.s.post / rchisq(n=1,df = kappa0.nu.post)
    struct$kappap[i+1] = GIGrvg::rgig(n=1, kappap.lambda.post, kappap.chi.post, kappap.psi.post)
  }
  }
  struct$kappa0 = struct$kappa0[(parameters$S.burnin+1):(parameters$S.burnin + parameters$S)]
  struct$kappap = struct$kappap[(parameters$S.burnin+1):(parameters$S.burnin + parameters$S)]
  struct$B0.posterior = struct$B0.posterior[,,(parameters$S.burnin+1):(parameters$S.burnin + parameters$S)]
  struct$Bp.posterior = struct$Bp.posterior[,,(parameters$S.burnin+1):(parameters$S.burnin + parameters$S)]
  
  # Set normalising matrix
  B0.last_draw = struct$B0.posterior[,,parameters$S]
  B0_hat = diag(sign(diag(B0.last_draw))) %*% B0.last_draw
  
  # Compute normalised B0
  B0.norm = normalize.Gibbs.output.parallel(struct$B0.posterior, B0_hat)
  
  # Compute and save normalised BP
  for (i in 1:parameters$S){
    struct$Bp.posterior[,,i] = B0.norm[,,i] %*% solve(struct$B0.posterior[,,i]) %*% struct$Bp.posterior[,,i]
  }
  
  # Save normalised B0
  struct$B0.posterior = B0.norm
  
  return(struct)
  
}

```

```{r  function to compute struct posterior means, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: Compute Posterior Means Function"
structPosteriorMeans = function(struct){
  struct.means = list(
  kappa0 = mean(struct$kappa0),
  kappap = mean(struct$kappap),
  B0 = rowMeans(struct$B0.posterior, dims = 2),
  Bp = rowMeans(struct$Bp.posterior, dims = 2)
  )
}
```

### Simulation Run

***Step 1:*** The following code computes the prior, initialises matrices to store results, and samples the draws for the simulated data.
```{r  Running extension for simulated data, echo=TRUE, eval = FALSE}
ext.sim.prior = ext.priors(parameters, artificialdata)
ext.sim.struct = init.struct(artificialdata, parameters$S + parameters$S.burnin)
ext.sim.struct <- ext.sampling(parameters, ext.sim.struct, ext.sim.prior, artificialdata)
```

```{r  save simulation struct, eval=FALSE}
save(ext.sim.struct, file = "ext.sim.struct.RData")
```

```{r  load simulation struct, eval=TRUE}
load(file = "ext.sim.struct.RData")
```

***Step 2:*** The following code computes and stores the posterior means of $\{\kappa_0^{(s)}, \kappa_+^{(s)}, B_0^{(s)}, B_+^{(s)}\}_{s =1}^{S}$.
```{r  compute the simulated struct posterior means for the data, echo=TRUE}
ext.sim.struct.means = structPosteriorMeans(ext.sim.struct)
```

***Step 3:*** The code snippets below display the sample posterior means of the estimated objects using simulated data.

The following code computes the posterior mean of $B_0$.
```{r  display in latex mode the simulated posterior means of B_0, results='asis', echo=TRUE}
ext.sim.data_B0 = pmatrix(ext.sim.struct.means$B0)
```
We can see that the posterior $B_0$ matrix is numerically equal to an identity matrix.

The following code computes the posterior mean of $B_+$.
```{r display sample average B_+, results='asis', echo=TRUE}
ext.sim.data_Bp = pmatrix(ext.sim.struct.means$Bp)

```
Here, as well, we can see that the first column of the autoregressive matrix is close to zero, while the rest of the matrix is numerically close to an identity matrix. Thus, we can see that the sampler works as the posterior draws replicate the data-generating process.

The following code computes the posterior mean of $\kappa_0$.
```{r display sample average draw of kappa_0, results='asis', echo=TRUE}
ext.sim.data_kappa0 = pmatrix(ext.sim.struct.means$kappa0)
```

The following code computes the posterior mean of $\kappa_+$.
```{r display sample average draw of kappa_+, results='asis', echo=TRUE}
ext.sim.data_kappap = pmatrix(ext.sim.struct.means$kappap)

```

The following graphs plot the histogram of kappa draws for the simulated data. The draws are concentrated towards zero with fat tails with $\kappa_0$ dispersed much more than $\kappa_+$.

```{r ext histogram of kappa draws simulated data}
par(mfrow=c(1,2))
hist(ext.sim.struct$kappa0,
     breaks = 100,
     col = "deepskyblue2",
     border = 'black',
     main = expression(paste("Distribution of ", kappa[0], " draws")),
     xlab = ""
     )
hist(ext.sim.struct$kappap,
     breaks = 100,
     col = 'deepskyblue2',
     border = 'black',
     main = expression(paste("Distribution of ", kappa["+"], " draws")),
     xlab = ""
     )
```

### Data Results

***Step 1:*** The following code computes the prior, initialises matrices to store results, and samples the draws for the actual data.
```{r  Running extension for actual data, eval=FALSE, echo=TRUE}
ext.prior  = ext.priors(parameters, mydata)
ext.struct = init.struct(mydata, parameters$S + parameters$S.burnin)
ext.struct = ext.sampling(parameters, ext.struct, ext.prior, mydata)
```

```{r  save struct for extension model, eval=FALSE}
save(ext.struct, file = "ext.struct.RData")
```

```{r  load struct for extension model, eval=TRUE}
load(file = "ext.struct.RData")
```



***Step 2:*** The following code computes and stores the posterior means of $\{\kappa_0^{(s)}, \kappa_+^{(s)}, B_0^{(s)}, B_+^{(s)}\}_{s =1}^{S}$.
```{r  compute the struct posterior means for the data, echo=TRUE}
ext.struct.means = structPosteriorMeans(ext.struct)
```

***Step 3:*** The code snippets below display the sample posterior means of the estimated objects using actual data.

```{r  display in latex mode the posterior means matrices, results='asis', echo=TRUE}
ext.data_B0 = pmatrix(ext.struct.means$B0)

```

```{r display the extension model posterior mean of B_0, results='asis', echo=TRUE}
ext.data_Bp = pmatrix(ext.struct.means$Bp)
```

```{r display the extension model posterior mean of kappa_0, results='asis', echo=TRUE}
ext.data_kappa0 = pmatrix(ext.struct.means$kappa0)
```

```{r display the extension model posterior mean of kappa_+, results='asis', echo=TRUE}
ext.data_kappap = pmatrix(ext.struct.means$kappap)
```

I plot the histogram of posterior kappa draws. $\kappa_0$ is drawn from an $\mathcal{IG}2$ distribution while $\kappa_+$ is drawn from a $\mathcal{GIG}$ distribution. Both draws are positively skewed with fat tails, with $\kappa_0$ having orders of magnitude higher dispersion. 
```{r ext histogram of kappa draws}
par(mfrow=c(1,2))
hist(ext.struct$kappa0,
     breaks = 100,
     col = "deepskyblue2",
     border = 'black',
     main = expression(paste("Distribution of ", kappa[0], " draws")),
     xlab = ""
     )
hist(ext.struct$kappap,
     breaks = 100,
     col = 'deepskyblue2',
     border = 'black',
     main = expression(paste("Distribution of ", kappa["+"], " draws")),
     xlab = ""
     )
```

## Extension 2: Inverse Normal Gamma Prior Hyperparameter Estimation
### Prior Distribution
Alternatively, I estimate $(\kappa_0, \kappa_+)$ assuming that these shrinkage parameters follow an inverse gamma 2 $(\mathcal{IG}2)$ distribution.
$$
\begin{align*}
    \kappa_0 | \underline s_{\kappa_0}, \underline \nu_{\kappa_0} \sim \mathcal{IG}2(\underline s_{\kappa_0}, \underline \nu_{\kappa_0})  && \kappa_+ | \underline s_{\kappa_+}, \underline \nu_{\kappa_+} \sim \mathcal{IG2}(\underline s_{\kappa_+},\underline \nu_{\kappa_+}) 
  \end{align*}
$$
The `ext.priors` function can be used to initialise these priors as well.

### Posterior Distribution
Thus, posteriors can be written as
$$
  \begin{align*}
    & p(\kappa_0 | Y, X, B_0) \propto p(B_0|\kappa_0) p(\kappa_0 | \underline s_{\kappa_0}, \underline \nu_{\kappa_0} )\\
    & \propto \prod_{n=1}^{N}\left(\kappa_0^{-\frac{1}{2}}\right)^{r_n} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N b_n V_n (\kappa_0 I_N)^{-1} V_n' b_n'  \right\} \cdot \kappa_0^{-\frac{\underline \nu_{\kappa_0}+2}{2}} \exp \left\{  -\frac{\underline s_{\kappa_0}}{2 \kappa_0} \right\}\\ 
    & \propto \kappa_0^{ -\frac{\underline \nu_{\kappa_0} + 2 + \sum_{n=1}^{N}r_n}{2}} \cdot \exp \left\{  -\frac{1}{2 \kappa_0} \sum_{n=1}^N b_n V_n V_n' b_n' + \underline s_{\kappa_0} \right\}\\
  \end{align*}
$$
  This gives
$$\bar s_{\kappa_0} =  \sum_{n=1}^N b_n V_n V_n' b_n' + \underline s_{\kappa_0}.$$
$$\bar \nu_{\kappa_0} = \underline \nu_{\kappa_0} +\sum_{n=1}^{N}r_n$$
  Also,
$$
  \begin{align*}
    & p(\kappa_+ | Y, X, B_+, B_0, \kappa_0) \propto p(B_+|\kappa_+, B_0, \kappa_0) p(\kappa_+ | \underline s_{\kappa_+}, \underline \nu_{\kappa_+} )\\
    & \propto \kappa_+^{-\frac{KN}{2}} \exp \left\{  -\frac{1}{2} \sum_{n=1}^N (B_{+n} - b_n V_n \underline B) (\kappa_+ I_{K})^{-1} (B_{+n} - b_n V_n \underline B)'  \right\} \cdot \kappa_+^{-\frac{\underline \nu_{\kappa_+} + 2}{2}} \exp \left\{-\frac{\underline s_{\kappa_+}}{2\kappa_+} \right\}\\
    & = \kappa_+^{-\frac{\underline \nu_{\kappa_+} + KN +2}{2}} \cdot \exp \left\{  -\frac{1}{2\kappa_+} \left(\sum_{n=1}^N (B_{+n} - b_n V_n \underline B)  (B_{+n} - b_n V_n \underline B)' + \underline s_{\kappa_+}   \right) \right\}\\
  \end{align*}
$$
  This gives
$$\bar s_{\kappa_+} = \sum_{n=1}^N (B_{+n} - b_n V_n \underline B)  (B_{+n} - b_n V_n \underline B)'  + \underline s_{\kappa_+}.$$
$$\bar \nu_{\kappa_+} = \underline \nu_{\kappa_+} +KN$$

### (Gibbs) Sampling
Given the full conditional posteriors of both $(\kappa_0,\kappa_+)$ and $(B_0, B_+)$, we can implement the Gibbs sampler as follows:

- Step 1: Initialise $(\kappa_0^{(0)},\kappa_+^{(0)})$. 

- Step 2: Use $(\kappa_0^{(s-1)},\kappa_+^{(s-1)})$ values to compute posterior parameters for each draw $(\bar B, \bar \Omega, \bar s, \bar \nu)^{(s)}$.

- Step 3: draw $(B_0, B_+)^{(s)} \sim \mathcal{NGN}(\bar B, \bar \Omega, \bar s, \bar \nu)^{(s)}$ using the sampling procedure presented in the baseline model.

- Step 4: draw $\kappa_0^{(s)} \sim p(\kappa_0 | Y, X, B_0^{(s)}) = \mathcal{IG2}(\bar s_{\kappa_0}, \bar \nu_{\kappa_0})$ and $\kappa_+^{(s)} \sim p(\kappa_+ | Y, X, B_0^{(s)}, B_+^{(s)}) = \mathcal{IG2}(\bar s_{\kappa_+}, \bar \nu_{\kappa_+})$.

Repeat steps 2-4 to get desired sample size and burn in some initial sample observations as needed.

The sampling function `ext2.sampling` is given below. 
```{r  Extension 2 Sampling, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: Sampler for Extension 2"
ext2.sampling = function(parameters, struct, priors, usedata){
  set.seed(12345)
  B0Vlist.initial = ltexclusion(usedata)
  B0.initial = B0Vlist.initial$B0
  for (i in 1:(parameters$S + parameters$S.burnin)){
    
    # Computing posterior parameters for each draw
    Omega.inv      = solve(struct$kappap[i] * priors$Omega)
    Omega.post.inv = usedata$X%*%t(usedata$X) + Omega.inv
    Omega.post     = solve(Omega.post.inv)
    B.post         = (usedata$Y%*%t(usedata$X) + priors$B%*%Omega.inv) %*% Omega.post
    S.post         = usedata$Y%*%t(usedata$Y) + solve(struct$kappa0[i] * priors$S) + priors$B%*%Omega.inv%*%t(priors$B) - B.post%*%Omega.post.inv%*%t(B.post)
    nu.post        = ncol(usedata$Y) + priors$nu


  if (i > 1){
    B0.initial = struct$B0.post[,,i-1]
  }
  
  B0.i = rgn(n=1, S.inv = S.post, nu = nu.post, V = B0Vlist.initial$V, B0.initial = B0.initial)
  # B0.hat = t(chol((nu.post - usedata$N)*S.post))
  # B0.norm.i = normalize.Gibbs.output.parallel(B0.i, B0.hat)
  Bp.i = rnorm.ngn(B0.i, B.post, Omega.post)
  
  struct$B0.posterior[,,i] = B0.i
  struct$Bp.posterior[,,i] = Bp.i
  
  # Now, update kappa
  kappa0.nu.post = priors$kappa0.nu + (usedata$N/2)
  kappap.nu.post = priors$kappap.nu + usedata$K * usedata$N
  kappa0.s.post  = priors$kappa0.s
  kappap.s.post  = priors$kappap.s
  # kappap.chi.post = 0
  for (n in 1:usedata$N){
    kappa0.s.post = kappa0.s.post + sum(B0.i[n,,1]^2)
    kappap.s.post = kappap.s.post + (Bp.i[n,,1] - B0.i[n,,1] %*% priors$B)%*%t(Bp.i[n,,1] - B0.i[n,,1] %*% priors$B)
  }
  
  if (i != (parameters$S + parameters$S.burnin)){
    struct$kappa0[i+1] = kappa0.s.post / rchisq(n=1,df = kappa0.nu.post)
    struct$kappap[i+1] = kappap.s.post / rchisq(n=1,df = kappap.nu.post)
  }
  }
  struct$kappa0 = struct$kappa0[(parameters$S.burnin+1):(parameters$S.burnin + parameters$S)]
  struct$kappap = struct$kappap[(parameters$S.burnin+1):(parameters$S.burnin + parameters$S)]
  struct$B0.posterior = struct$B0.posterior[,,(parameters$S.burnin+1):(parameters$S.burnin + parameters$S)]
  struct$Bp.posterior = struct$Bp.posterior[,,(parameters$S.burnin+1):(parameters$S.burnin + parameters$S)]
  
  # Set normalising matrix
  B0.last_draw = struct$B0.posterior[,,parameters$S]
  B0_hat = diag(sign(diag(B0.last_draw))) %*% B0.last_draw
  
  # Compute normalised B0
  B0.norm = normalize.Gibbs.output.parallel(struct$B0.posterior, B0_hat)
  
  # Compute normalised BP
  for (i in 1:parameters$S){
    struct$Bp.posterior[,,i] = B0.norm[,,i] %*% solve(struct$B0.posterior[,,i]) %*% struct$Bp.posterior[,,i]
  }
  
  # Save normalised B0
  struct$B0.posterior = B0.norm
  
  return(struct)
  
}

```

### Simulation Run
***Step 1:*** The following code computes the prior, initialises matrices to store results, and samples the draws for the simulated data.
```{r  Running extension 2 for simulated data, echo=TRUE, eval = FALSE}
ext2.sim.prior = ext.priors(parameters, artificialdata)
ext2.sim.struct = init.struct(artificialdata, parameters$S + parameters$S.burnin)
ext2.sim.struct <- ext2.sampling(parameters, ext2.sim.struct, ext2.sim.prior, artificialdata)
```

```{r  save simulation 2 struct, eval=FALSE}
save(ext2.sim.struct, file = "ext2.sim.struct.RData")
```

```{r  load simulation 2 struct, eval=TRUE}
load(file = "ext2.sim.struct.RData")
```

***Step 2:*** The following code computes and stores the posterior means of $\{\kappa_0^{(s)}, \kappa_+^{(s)}, B_0^{(s)}, B_+^{(s)}\}_{s =1}^{S}$.
```{r  compute the simulated 2 struct posterior means for the data, echo=TRUE}
ext2.sim.struct.means = structPosteriorMeans(ext2.sim.struct)
```

***Step 3:*** The code snippets below display the sample posterior means of the estimated objects using simulated data.

The following code computes the posterior mean of $B_0$.
```{r  display in latex mode the simulated 2 posterior means of B_0, results='asis', echo=TRUE}
ext2.sim.data_B0 = pmatrix(ext2.sim.struct.means$B0)
```

The following code computes the posterior mean of $B_+$.
```{r ext2 display sample average B_+, results='asis', echo=TRUE}
ext2.sim.data_Bp = pmatrix(ext2.sim.struct.means$Bp)

```

Again, we can see that the first column of the autoregressive matrix is close to zero, while the rest of the $B_+$ matrix and the $B_0$ matrix both are numerically close to an identity matrix. Thus, we can see that this sampler also works since the posterior draws replicate the data-generating process.

The following code computes the posterior mean of $\kappa_0$.
```{r ext2 display sample average draw of kappa_0, results='asis', echo=TRUE}
ext2.sim.data_kappa0 = pmatrix(ext2.sim.struct.means$kappa0)
```

The following code computes the posterior mean of $\kappa_+$.
```{r ext2 display sample average draw of kappa_+, results='asis', echo=TRUE}
ext2.sim.data_kappap = pmatrix(ext2.sim.struct.means$kappap)

```

I plot the histogram of posterior kappa draws. both $\kappa_0$ and $\kappa_+$ are drawn from an $\mathcal{IG}2$ distribution. $\kappa_0$ draws are more positively skewed while $\kappa_+$ draws have fatter tails, with $\kappa_0$ having orders of magnitude higher dispersion. 

```{r ext2 histogram of kappa draws simulated data}
par(mfrow=c(1,2))
hist(ext2.sim.struct$kappa0,
     breaks = 100,
     col = "deepskyblue2",
     border = 'black',
     main = expression(paste("Distribution of ", kappa[0], " draws")),
     xlab = ""
     )
hist(ext2.sim.struct$kappap,
     breaks = 100,
     col = 'deepskyblue2',
     border = 'black',
     main = expression(paste("Distribution of ", kappa["+"], " draws")),
     xlab = ""
     )
```

### Data Results

***Step 1:*** The following code computes the prior, initialises matrices to store results, and samples the draws for the actual data.
```{r  ext2 Running extension for actual data, eval = FALSE, echo=TRUE}
ext2.prior  = ext.priors(parameters, mydata)
ext2.struct = init.struct(mydata, parameters$S + parameters$S.burnin)
ext2.struct = ext2.sampling(parameters, ext2.struct, ext2.prior, mydata)
```

```{r  ext2 save struct for extension model, eval=FALSE}
save(ext2.struct, file = "ext2.struct.RData")
```

```{r  ext2 load struct for extension model, eval=TRUE}
load(file = "ext2.struct.RData")
```


***Step 2:*** The following code computes and stores the posterior means of $\{\kappa_0^{(s)}, \kappa_+^{(s)}, B_0^{(s)}, B_+^{(s)}\}_{s =1}^{S}$.
```{r  ext2 compute the struct posterior means for the data, echo=TRUE}
ext2.struct.means = structPosteriorMeans(ext2.struct)
```

***Step 3:*** The code snippets below display the sample posterior means of the estimated objects using actual data.
```{r  ext2 display in latex mode the posterior means matrices, results='asis', echo=TRUE}
ext2.data_B0 = pmatrix(ext2.struct.means$B0)

```

```{r ext2 display the extension model posterior mean of B_0, results='asis', echo=TRUE}
ext2.data_Bp = pmatrix(ext2.struct.means$Bp)
```

```{r ext2 display the extension model posterior mean of kappa_0, results='asis', echo=TRUE}
ext2.data_kappa0 = pmatrix(ext2.struct.means$kappa0)
```

```{r ext2 display the extension model posterior mean of kappa_+, results='asis', echo=TRUE}
ext2.data_kappap = pmatrix(ext2.struct.means$kappap)
```


```{r ext2 histogram of kappa draws}
par(mfrow=c(1,2))
hist(ext2.struct$kappa0,
     breaks = 100,
     col = "deepskyblue2",
     border = 'black',
     main = expression(paste("Distribution of ", kappa[0], " draws")),
     xlab = ""
     )
hist(ext2.struct$kappap,
     breaks = 100,
     col = 'deepskyblue2',
     border = 'black',
     main = expression(paste("Distribution of ", kappa["+"], " draws")),
     xlab = ""
     )
```

## Extension 3: Stochastic Volatility
Next, I introduce stochastic volatility in the form of common conditional heteroskedasticity of the structural errors. We assume that
$$
u_t|X \sim N_N(0_N, diag(\sigma^2_t I_n))
$$
where $\sigma^2 = (\exp(h_1), \dots, \exp(h_T))$ is the $T-$vector. This introduction of stochastic volatility transforms the posterior parameters of $(B_+,B_0)$ as follows:
$$
\begin{align*}
&\bar \Omega = (X diag(\sigma^2) X' + \underline \Omega^{-1})^{-1}\\
&\bar B = (Ydiag(\sigma^2)X' + \underline B \underline \Omega^{-1}) \bar \Omega\\
&\bar S = (Ydiag(\sigma^2)Y' + \underline S^{-1} + \underline B \underline \Omega^{-1} \underline B' - \bar B \bar \Omega^{-1} \bar B')^{-1}
\end{align*}
$$
while $\bar \nu = T + \underline \nu$ updates the same way as before.

### (Gibbs) Sampling

```{r SV priors}
SVpriors= function(T){
  H = diag(T) - rbind(rep(0,1,T),diag(1,T-1,T))
  priors = list(
  h0.v = 1, 
  h0.m = 0,
  sigmav.s = 0.01,
  sigmav.nu = 10,
  HH = t(H)%*%H
)
}
```

```{r sv aux}
SVaux = function(Y,X, B0, Bp, T){
  aux = list(
    Y = Y,
    X = X,
    H = array(0,c(T,1)),
    h0 = 0,
    sigma.v2 = 1,
    s = array(1,c(T,1)),
    Bplus = Bp,
    B0 = B0,
    sigma2 = array(1,c(T,1))
  )
}
```


The Gibbs sampler for the stochastic volatility component is implemented in the `SVcommonSF.Gibbs.iteration` function below.
```{r ext3 SV gibbs sampler, echo=TRUE}
#| code-fold: true
#| code-summary: "Common SV component sampler"
############################################################
# prepared by Tomasz Woźniak
############################################################

SVcommonSF.Gibbs.iteration = function(aux, priors){
  # A single iteration of the Gibbs sampler for the SV component
  #
  # aux is a list containing:
  #   Y - a NxT matrix
  #   X - a KxT matrix
  #   H - a Tx1 matrix
  #   h0 - a scalar
  #   sigma.v2 - a scalar
  #   s - a Tx1 matrix
  #   Bplus - a NxK matrix
  #   B0 - an NxN matrix
  #   sigma2 - a T-vector
  #
  # priors is a list containing:
  #   h0.v - a positive scalar
  #   h0.m - a scalar
  #   sigmav.s - a positive scalar
  #   sigmav.nu - a positive scalar
  #   HH - a TxT matrix
  
  T             = dim(aux$Y)[2]
  N             = dim(aux$Y)[1]
  alpha.st      = c(1.92677,1.34744,0.73504,0.02266,0-0.85173,-1.97278,-3.46788,-5.55246,-8.68384,-14.65000)
  sigma.st      = c(0.11265,0.17788,0.26768,0.40611,0.62699,0.98583,1.57469,2.54498,4.16591,7.33342)
  pi.st         = c(0.00609,0.04775,0.13057,0.20674,0.22715,0.18842,0.12047,0.05591,0.01575,0.00115)
  
  Z             = rowSums( t( aux$B0 %*% aux$Y - aux$Bplus %*% aux$X ) ) / sqrt(N)
  Y.tilde       = as.vector(log((Z + 0.0000001)^2))
  Ytilde.alpha  = as.matrix(Y.tilde - alpha.st[as.vector(aux$s)])
  
  # sampling initial condition
  ############################################################
  V.h0.bar      = 1/((1 / priors$h0.v) + (1 / aux$sigma.v2))
  m.h0.bar      = V.h0.bar*((priors$h0.m / priors$h0.v) + (aux$H[1] / aux$sigma.v2))
  h0.draw       = rnorm(1, mean = m.h0.bar, sd = sqrt(V.h0.bar))
  aux$h0        = h0.draw
  
  # sampling sigma.v2
  ############################################################
  sigma.v2.s    = priors$sigmav.s + sum(c(aux$H[1] - aux$h0, diff(aux$H))^2)
  sigma.v2.draw = sigma.v2.s / rchisq(1, priors$sigmav.nu + T)
  aux$sigma.v2  = sigma.v2.draw
  
  # sampling auxiliary states
  ############################################################
  Pr.tmp        = simplify2array(lapply(1:10,function(x){
    dnorm(Y.tilde, mean = as.vector(aux$H + alpha.st[x]), sd = sqrt(sigma.st[x]), log = TRUE) + log(pi.st[x])
  }))
  Pr            = t(apply(Pr.tmp, 1, function(x){exp(x - max(x)) / sum(exp(x - max(x)))}))
  s.cum         = t(apply(Pr, 1, cumsum))
  r             = matrix(rep(runif(T), 10), ncol = 10)
  ss            = apply(s.cum < r, 1, sum) + 1
  aux$s         = as.matrix(ss)
  
  
  # sampling log-volatilities using functions for tridiagonal precision matrix
  ############################################################
  Sigma.s.inv   = diag(1 / sigma.st[as.vector(aux$s)])
  D.inv         = Sigma.s.inv + (1 / aux$sigma.v2) * priors$HH
  b             = as.matrix(Ytilde.alpha / sigma.st[as.vector(aux$s)] + (aux$h0/aux$sigma.v2)*diag(T)[,1])
  lead.diag     = diag(D.inv)
  sub.diag      = mgcv::sdiag(D.inv, -1)
  D.chol        = mgcv::trichol(ld = lead.diag, sd = sub.diag)
  D.L           = diag(D.chol$ld)
  mgcv::sdiag(D.L,-1) = D.chol$sd
  x             = as.matrix(rnorm(T))
  a             = forwardsolve(D.L, b)
  draw          = backsolve(t(D.L), a + x)
  aux$H         = as.matrix(draw)
  aux$sigma2    = as.vector(exp(draw))

  return(aux)
}
```


```{r  ext3 sampling, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: Gibbs Sampler for Extension 3"
ext3.sampling = function(parameters, struct, priors, usedata){
  set.seed(12345)
  T = ncol((usedata$Y))
  # Step 0: initialize B_0 and B_+
  B0Vlist.initial     = ltexclusion(usedata)
  
  Omega.inv           = solve(priors$Omega)
  Omega.post.inv = usedata$X%*%t(usedata$X) + Omega.inv
  Omega.post     = solve(Omega.post.inv)
  B.post         = (usedata$Y%*%t(usedata$X) + priors$B%*%Omega.inv) %*% Omega.post
  nu.post        = ncol(usedata$Y) + priors$nu
  
  for (i in 1:(parameters$S + parameters$S.burnin)){
  
  # Set in-loop definition of B0.post and Bp.post
  if (i == 1){
    B0.post = B0Vlist.initial$B0
    Bp.post = cbind(array(0,c(N,1)), array(diag(N),c(usedata$N, usedata$N * usedata$p))) 
  } else {
    B0.post = struct$B0.posterior[,,i-1]
    Bp.post = struct$Bp.posterior[,,i-1]
  }
  
  # Step 1a: Now, update kappa0 and kappa_+ posteriors
  kappa0.nu.post = priors$kappa0.nu + (usedata$N/2)
  kappap.lambda.post = -(-priors$kappap.nu + usedata$K * usedata$N)/2
  kappap.psi.post = 1/priors$kappap.s
  kappa0.s.post = priors$kappa0.s
  for (n in 1:usedata$N){
    kappa0.s.post = kappa0.s.post + sum(B0.post^2)
    kappap.chi.post = (Bp.post - B0.post %*% B.post) %*% t(Bp.post - B0.post %*% B.post)
  }
  
  # Step 1b: draw kappa0 from IG2 and kappa_+ from GIG distributions
  struct$kappa0[i] = kappa0.s.post / rchisq(n=1,df = kappa0.nu.post)
  struct$kappap[i] = GIGrvg::rgig(n=1, kappap.lambda.post, kappap.chi.post, kappap.psi.post)
  
  #Step 2a: initialise inputs aux and prior
  aux = SVaux(usedata$Y, usedata$X, B0.post, Bp.post, T)
  SVprior = SVpriors(T)
  
  # Step 2b: draw the common heteroskedastic sigma2
  aux = SVcommonSF.Gibbs.iteration(aux,SVprior)
  sigma2 = aux$sigma2
  
  # Step 3a: compute posterior parameters for B0, B_+
  Omega.inv      = solve(struct$kappap[i] * priors$Omega)
  Omega.post.inv = usedata$X%*% diag(sigma2) %*%t(usedata$X) + Omega.inv
  Omega.post     = solve(Omega.post.inv)
  B.post         = (usedata$Y%*%diag(sigma2) %*%t(usedata$X) + priors$B%*%Omega.inv) %*% Omega.post
  S.inv.post     = usedata$Y%*%diag(sigma2) %*%t(usedata$Y) + solve(struct$kappa0[i] * priors$S) + priors$B%*%Omega.inv%*%t(priors$B) - B.post%*%Omega.post.inv%*%t(B.post)
  
  # Step 3b: draw non-normalised B0 and B_+

  B0.i = rgn(n=1, S.inv = S.inv.post, nu = nu.post, V = B0Vlist.initial$V, B0.initial = B0.post)
  Bp.i = rnorm.ngn(B0.i, B.post, Omega.post)
  
  struct$B0.posterior[,,i] = B0.i
  struct$Bp.posterior[,,i] = Bp.i
  }

  # Discard the burn-in results
  struct$kappa0 = struct$kappa0[(parameters$S.burnin+1):(parameters$S.burnin + parameters$S)]
  struct$kappap = struct$kappap[(parameters$S.burnin+1):(parameters$S.burnin + parameters$S)]
  
  struct$B0.posterior = struct$B0.posterior[,,(parameters$S.burnin+1):(parameters$S.burnin + parameters$S)]
  struct$Bp.posterior = struct$Bp.posterior[,,(parameters$S.burnin+1):(parameters$S.burnin + parameters$S)]
  
  # Set normalising matrix
  B0.last_draw = struct$B0.posterior[,,parameters$S]
  B0_hat = diag(sign(diag(B0.last_draw))) %*% B0.last_draw
  
  # Compute normalised B0
  B0.norm = normalize.Gibbs.output.parallel(struct$B0.posterior, B0_hat)
  
  # Compute normalised BP
  for (i in 1:parameters$S){
    struct$Bp.posterior[,,i] = B0.norm[,,i] %*% solve(struct$B0.posterior[,,i]) %*% struct$Bp.posterior[,,i]
  }
  
  # Save normalised B0
  struct$B0.posterior = B0.norm
  return(struct)
  
}

```

### Data Results

***Step 1:*** The following code computes the prior, initialises matrices to store results, and samples the draws for the actual data.
```{r  ext3 Running extension for actual data, eval = FALSE, include=TRUE}
#| code-fold: true
#| code-summary: "View code"
ext3.prior = ext.priors(parameters, mydata)
ext3.struct = init.struct(mydata, parameters$S + parameters$S.burnin)
ext3.struct <- ext3.sampling(parameters, ext3.struct, ext3.prior, mydata)
```

```{r  ext3 save struct for extension model, eval=FALSE}
save(ext3.struct, file = "ext3.struct.RData")
```

```{r  ext3 load struct for extension model, eval=TRUE}
load(file = "ext3.struct.RData")
```



***Step 2:*** The following code computes and stores the posterior means of $\{\kappa_0^{(s)}, \kappa_+^{(s)}, B_0^{(s)}, B_+^{(s)}\}_{s =1}^{S}$.
```{r  ext3 compute the struct posterior means for the data, echo=TRUE}
ext3.struct.means = structPosteriorMeans(ext3.struct)
```

***Step 3:*** The code snippets below display the sample posterior means of the estimated objects using actual data.

```{r  ext3 display in latex mode the posterior means matrices, results='asis', echo=TRUE}
ext3.data_B0 = pmatrix(ext3.struct.means$B0)

```

```{r ext3 display the extension model posterior mean of B_0, results='asis', echo=TRUE}
ext3.data_Bp = pmatrix(ext3.struct.means$Bp)
```

```{r ext3 display the extension model posterior mean of kappa_0, results='asis', echo=TRUE}
ext3.data_kappa0 = pmatrix(ext3.struct.means$kappa0)
```

```{r ext3 display the extension model posterior mean of kappa_+, results='asis', echo=TRUE}
ext3.data_kappap = pmatrix(ext3.struct.means$kappap)
```

```{r ext3 histogram of kappa draws}
par(mfrow=c(1,2))
hist(ext3.struct$kappa0,
     breaks = 100,
     col = "deepskyblue2",
     border = 'black',
     main = expression(paste("Distribution of ", kappa[0], " draws")),
     xlab = ""
     )
hist(ext3.struct$kappap,
     breaks = 100,
     col = 'deepskyblue2',
     border = 'black',
     main = expression(paste("Distribution of ", kappa["+"], " draws")),
     xlab = ""
     )
```

## Analysis of Posteriors
The following graph plots the histogram of posterior $(\kappa_0, \kappa_+)$ draws for each of the model extensions implemented above.
```{r histogram of kappa draws all extensions}
par(mfrow=c(3,2), mar=c(2,2,2,2))
hist(ext.struct$kappa0,
     breaks = 100,
     col = "deepskyblue2",
     border = 'black',
     main = expression(paste("Extension 1: Distribution of ", kappa[0], " draws")),
     xlab = ""
     )
hist(ext.struct$kappap,
     breaks = 100,
     col = 'deepskyblue2',
     border = 'black',
     main = expression(paste("Extension 1: Distribution of ", kappa["+"], " draws")),
     xlab = ""
     )

hist(ext2.struct$kappa0,
     breaks = 100,
     col = "deepskyblue2",
     border = 'black',
     main = expression(paste("Extension 2: Distribution of ", kappa[0], " draws")),
     xlab = ""
     )
hist(ext2.struct$kappap,
     breaks = 100,
     col = 'deepskyblue2',
     border = 'black',
     main = expression(paste("Extension 2: Distribution of ", kappa["+"], " draws")),
     xlab = ""
     )

hist(ext3.struct$kappa0,
     breaks = 100,
     col = "deepskyblue2",
     border = 'black',
     main = expression(paste("Extension 3: Distribution of ", kappa[0], " draws")),
     xlab = ""
     )
hist(ext3.struct$kappap,
     breaks = 100,
     col = 'deepskyblue2',
     border = 'black',
     main = expression(paste("Extension 3: Distribution of ", kappa["+"], " draws")),
     xlab = ""
     )
```

Extension 3 posits the same priors as Extension 1, but is enriched by inclusion of common stochastic volatility component in the structural shocks. As a result, we see that the $\kappa_0$ draws become more positively skewed with even higher dispersion. However, the effect is reverse for the $\kappa_+$ draws. In extension 3, these draws are more dispersed and exhibit fatter tails. 

Interestingly, the $\kappa_+$ draws under the $\mathcal{IG}2$ prior lead to posterior draws that exhibit a bell curve structure with much lower positive skewness that than in other extensions. Moreover, the mode is also around 3 orders of magnitude higher. Predictably such difference in draws lead to much different looking impulse response functions shown in the next section.

# Empirical Investigation
In this section, I explore how the variables in my data respond to a domestic monetary policy shock (i.e. a 25 basis points cash rate shock) and a foreign monetary policy shock (i.e. a 25 basis points US Federal Funds Rate shock) using Impulse Response Function (IRF) graphs. I also plot the Forecast Error Variance Decomposition (FEVD) plots to illustrate what extent of the variability of my variables are explained by the structural shocks the model, and how these explainability evolves over time.

The `irf_fevd` function below computes and stores the posterior IRF and FEVD values for each variable and each shock. The implementation follows closely from [Tomasz Woźniak](https://github.com/donotdespair)'s lecture notes.

```{r  function to compute the irf and fevd matrices, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: Compute IRF/FEVD Function"
irf_fevd = function(parameters, usedata, Bposteriors){
  # Unpack variables
  N = usedata$N
  K = usedata$K
  p = usedata$p
  S = parameters$S
  h = parameters$h
  B0.posterior = Bposteriors$B0.posterior
  Bp.posterior = Bposteriors$Bp.posterior
  
  B.posterior       = array(NA,c(N,N,S))
  A.posterior       = array(NA,c(N,K,S))
for (s in 1:S){
  B                 = solve(B0.posterior[,,s])
  B.posterior[,,s]  = B
  A.posterior[,,s]  = B %*% Bp.posterior[,,s]
}

IRF.posterior       = array(NA,c(N,N,h+1,S))
IRF.inf.posterior   = array(NA,c(N,N,S))
FEVD.posterior      = array(NA,c(N,N,h+1,S))
J                   = cbind(diag(N),matrix(0,N,N*(p-1)))
for (s in 1:S){
  A.bold          = rbind(A.posterior[,2:(1+N*p),s],cbind(diag(N*(p-1)),matrix(0,N*(p-1),N)))
  IRF.inf.posterior[,,s]          = J %*% solve(diag(N*p)-A.bold) %*% t(J) %*% B.posterior[,,s]
  A.bold.power    = A.bold
  for (i in 1:(h+1)){
    if (i==1){
      IRF.posterior[,,i,s]        = B.posterior[,,s]
    } else {
      IRF.posterior[,,i,s]        = J %*% A.bold.power %*% t(J) %*% B.posterior[,,s]
      A.bold.power                = A.bold.power %*% A.bold
    }
    for (n in 1:N){
      for (nn in 1:N){
        FEVD.posterior[n,nn,i,s]  = sum(IRF.posterior[n,nn,1:i,s]^2)
      }
    }
    FEVD.posterior[,,i,s]         = diag(1/apply(FEVD.posterior[,,i,s],1,sum))%*%FEVD.posterior[,,i,s]
  }
}
FEVD.posterior    = 100*FEVD.posterior

list(IRF.posterior = IRF.posterior, IRF.inf.posterior= IRF.inf.posterior, FEVD.posterior = FEVD.posterior)
}
```

The next two subsections plot and explain the IRFs and selected FEVDs for the baseline and model extensions respectively.

## Baseline Model
```{r compute the irf and fevd results for data, eval=FALSE}
data.irf_fevd = irf_fevd(parameters, mydata, Bposteriors)
```

```{r save the irf and fevd results for data, eval=FALSE}
save(data.irf_fevd, file = "data.irf_fevd.RData")
```

```{r load the irf and fevd results for data, eval=TRUE}
load("data.irf_fevd.RData")
```

### Impulse Response Functions (IRFs)

The `plot_irf` function plots the IRF of all variables given the position of the structural shock. 
```{r  function that plots the irf of chosen variable position, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: Plot IRF Function"
plot_irf = function(position, fevd, parameters){
  # Unpack results
IRF.posterior = fevd$IRF.posterior

IRF.posterior.mps = IRF.posterior[,position,,]
IRFs.k1           = apply(IRF.posterior.mps,1:2,median)
IRF.posterior.mps = IRF.posterior.mps*(0.25/IRFs.k1[position,1])
IRFs.k1           = apply(IRF.posterior.mps,1:2,median)
IRFs.inf.k1       = apply(IRF.posterior.mps,1,mean)
var_names         = colnames(df)
rownames(IRFs.k1) = var_names

IRFs.k1.hdi       = apply(IRF.posterior.mps,1:2, hdi, credMass=0.68)
h                 = parameters$h
hh                = 1:(h+1)

# Define colours
mcxs1  = "#05386B"
mcxs2  = "#379683"
mcxs3  = "#5CDB95"
mcxs4  = "#8EE4AF"
mcxs5  = "#EDF5E1"
purple = "#b02442"

mcxs1.rgb   = col2rgb(mcxs1)
mcxs1.shade1= rgb(mcxs1.rgb[1],mcxs1.rgb[2],mcxs1.rgb[3], alpha=120, maxColorValue=255)
mcxs2.rgb   = col2rgb(mcxs2)
mcxs2.shade1= rgb(mcxs2.rgb[1],mcxs2.rgb[2],mcxs2.rgb[3], alpha=120, maxColorValue=255)

# Save as pdf
par(mfrow=c(3,3), mar=c(2,2,2,2))
for (n in 1:N){
  ylims     = range(IRFs.k1[n,hh],IRFs.k1.hdi[,n,1:4],0)
  p1 <- plot(hh,IRFs.k1[n,hh], type="l", ylim=ylims, axes=FALSE, xlab="", main=rownames(IRFs.k1)[n])

  axis(1,c(1,2,5,7,9,13,17),c("","1 quarter","1 year","6 quarters", "2 years", "3 years", "4 years"))
  
  axis(2,c(ylims[1],0,ylims[2]),round(c(ylims[1],0,ylims[2]),3))
  polygon(c(hh,(h+1):1), c(IRFs.k1.hdi[1,n,hh],IRFs.k1.hdi[2,n,(h+1):1]), col=mcxs1.shade1,border=mcxs1.shade1)
  abline(h=0)
  lines(hh, IRFs.k1[n,hh],lwd=2,col=mcxs1)
}
return(list(p1))
}
```

Using the `plot_irf` function, we plot the IRFs given a 25 basis points shock to the Australian Cash Rate below:
```{r irf of cashrate, echo=TRUE}
#| label: cashrate.irf
#| fig-cap: "IRF plots for the AU Cash Rate Shock"
#| fig-cap-location: top
cashrate.irf = plot_irf(5, data.irf_fevd, parameters)
```
The last row of plots contain housing variables. We can see that home loan rates move almost in line with the dynamics of the effect of positive domestic monetary policy shock on cash rate itself. 

The number of new homes built increases slightly at the time of the shock but continuously decreases after that reaching a trough in about 1.5 years. It trends upwards after that point but at a much lower rate than at which it dropped.

The housing price index starts from zero and trends downwards in a concave fashion over time stabilising at a lower level. This implies that a positive monetary shock reduces the House Price Index but that the magnitude of the negative effect increases but stabilises over time.


Now, we plot the IRFs given a +25 basis points shock to the US federal funds rate.
```{r irf of usffr, echo=TRUE}
#| label: usffr.irf
#| fig-cap: "IRF plots for the US Federal Funds Rate Shock"
#| fig-cap-location: top
usffr.irf = plot_irf(1, data.irf_fevd, parameters)
```

Here, we can see that a positive US Federal Funds rate shock affects the home loan rate in a similar way that it affects the RBA target cash rate. The impulse response for home loan rate starts at a small positive value and decreases very gradually over time. 

A positive US Federal Funds rate shock decreases the quantity of new homes. The negative effect peaks within a year and then roughly remains constant thereafter.

And lastly, a positive US Federal Funds rate shock increases the house price index and that effect roughly stays the same over time.

```{r save the irf figures as pdf, eval=FALSE}
pdf(file = "aucr_irf.pdf", onefile = FALSE, width = 16, height = 9)
plot_irf(8, data.irf_fevd, parameters)
dev.off()

pdf(file = "usffr_irf.pdf", onefile = FALSE, width = 16, height = 9)
plot_irf(4, data.irf_fevd, parameters)
dev.off()
```

### Forecast Error Variance Decomposition

The `plot_fevd` function plots the FEVD for a chosen variable using the `FEVD.posterior` matrices computed by the `irf_fevd` function. 
```{r function to plot the fevd of variables given position, echo=TRUE}
#| code-fold: true
#| code-summary: "Code: Plot FEVD Function"
plot_fevd = function(position, fevd, parameters){
  FEVD.posterior = fevd$FEVD.posterior
  h = parameters$h
  N = dim(FEVD.posterior)[1]
  ffr.pos = which(colnames(df) == "US FFR Max Target Rate")
  cr.pos = which(colnames(df) == "RBA Target Cash Rate")
  
  hh            = 1:(h+1)
fevd.mat  = apply(FEVD.posterior[position,,,],1:2,mean)
fevd.mat  = rbind(rep(0,h+1),apply(fevd.mat,2,cumsum))

colors = c("deepskyblue1","deepskyblue2","deepskyblue","deepskyblue3","deepskyblue4","dodgerblue",
           "maroon1","maroon","maroon2","magenta","maroon3","maroon4")

par(mar=rep(4,4),cex.axis=1, cex.lab=0.8)
p1 <- plot(hh,fevd.mat[1,], type="n", ylim=c(0,100), axes=FALSE, xlab="", ylab="")
axis(1,c(1,2,5,7,9,13,17),c("","1 quarter","1 year","6 quarters", "2 years", "3 years", "4 years"))
axis(2,c(0,50,100),c("",colnames(df)[position],""))
for (n in 1:N){
  polygon(c(hh,(h+1):1), c(fevd.mat[n,hh],fevd.mat[n+1,(h+1):1]), col=colors[n],border=colors[n])
}
axis(4, (0.5*(fevd.mat[(1:N),(h+1)]+fevd.mat[2:(N+1),(h+1)]))[c(ffr.pos,position,cr.pos)], c("us.mps","","au.mps")) # fix position of terms

return(list(p1))
}
```


```{r fevd for homeloan rates, warning=FALSE}
variable.choice = "Owner-occupied home loan rates"
position = which(colnames(df) == variable.choice)
hlr.fevd = plot_fevd(position, data.irf_fevd, parameters)
```
The FEVD graph above shows that the domestic monetary policy shock is very important in explaining the variability in home loan rates. Foreign monetary policy shocks roughly maintain the same level of explanability of the variation in home loan rates. It also indicates that an inflation shock would have an increasing explanatory power over time of home loan rates.


```{r fevd for quantity of new homes}
variable.choice = "Quantity of new homes (log)"
position = which(colnames(df) == variable.choice)
nh.fevd = plot_fevd(position, data.irf_fevd, parameters)
```

The FEVD graph above shows that both domestic and foreign monetary policy shock have a greater explanatory power on the variation of the quantity of new homes built over time. It indicates that the effects of these shocks have a long lasting effect and increasing effect on the quantity of homes built.

```{r fevd for house price index}
variable.choice = "House Price Index (log)"
position = which(colnames(df) == variable.choice)
hpi.fevd = plot_fevd(position, data.irf_fevd, parameters)
```
The FEVD graph above shows that both domestic and foreign monetary policy shock have little explanatory power over the variation in house price index at time $t$ but that this power increases over time. This also indicates that the effects of these shocks have a long lasting effect and increasing effect on house prices.

## Extension 1: Normal-Gamma Hyperparameter Prior
```{r compute the irf and fevd results for data in extended model, eval=FALSE}
ext.data.irf_fevd = irf_fevd(parameters, mydata, ext.struct)
```

```{r save the irf and fevd results for data in extended model, eval=FALSE}
save(ext.data.irf_fevd, file = "ext.data.irf_fevd.RData")
```

```{r load the irf and fevd results for data in extended model, eval=TRUE}
load("ext.data.irf_fevd.RData")
```

### Impulse Response Functions
For the extended model, using the `plot_irf` function, we plot the IRFs to a +25 basis points increase in the Australian Cash Rate.
```{r irf of cashrate extended model}
#| label: ext.cashrate.irf
#| fig-cap: "IRF plots for the AU Cash Rate Shock, Extended Model"
#| fig-cap-location: top
ext.cashrate.irf = plot_irf(5, ext.data.irf_fevd, parameters)
```
The IRF results are different compared to the baseline results. The effect of a positive domestic monetary policy shock on home loan rates seems to be more persistent as the effect erodes away at a much lower rate.

The quantity of new homes built is the qualitatively the same as under the baseline specification, except for the absence of an initial boost in the number of houses built. Quantitatively, the negative effect on quantity of new homes is much lower under this specification.

House price index results are also qualitatively the same as under the baseline specification. However, the magnitude of the reduction in house prices due to a positive monetary policy shock is also lower than under the baseline specification.


For the extended model, using the `plot_irf` function, we plot the IRFs to a +25 basis points increase in the US Federal Funds Rate.
```{r irf of usffr extended model}
#| label: ext.usffr.irf
#| fig-cap: "IRF plots for the US Federal Funds Rate Shock, Extended Model"
#| fig-cap-location: top
ext.usffr.irf = plot_irf(1, ext.data.irf_fevd, parameters)
```

The effect of a positive foreign monetary policy shock on home loan rates appears both qualitatively and quantitatively similar as under the baseline specification.

However, the effect of a positive foreign monetary policy shock on quantity of homes is different. There is an immediate positive increase in the number of new homes, but this effect linearly goes dissipates away returning to zero after around 3 years.

The effect on house price index is similar as under the baseline specification where a postive USFFR shock increases house price index slightly and keeps them at that level.

### Forecast Error Variance Decomposition

In the figures below, we plot the FEVDs for our selected housing variables of interest.

```{r fevd for homeloan rates extended model, warning=FALSE}
variable.choice = "Owner-occupied home loan rates"
position = which(colnames(df) == variable.choice)
ext.hlr.fevd = plot_fevd(position, ext.data.irf_fevd, parameters)
```

The FEVD graph above shows that the domestic monetary policy shock carries the bulk of the explanatory power in the variability of home loan rates. Compared to the baseline model, the shocks on real economy variables (GDP, consumption) are even less important (and carry virtually no explanatory power) while the change in CPI also become less important.

```{r fevd for quantity of new homes extended model}
variable.choice = "Quantity of new homes (log)"
position = which(colnames(df) == variable.choice)
ext.nh.fevd = plot_fevd(position, ext.data.irf_fevd, parameters)
```

The FEVD graph above shows that while both domestic and foreign monetary policy have little explanatory power on the variability of the quantity of new homes built, the power increases over time. However, this increase in explanatory power is lower than in the baseline model. Moreover, shocks on the real economic indicators (GDP, consumption) are relatively more important while change in inflation is less important in explaining the variability of the quantity of new homes.

```{r fevd for house price index extended model}
variable.choice = "House Price Index (log)"
position = which(colnames(df) == variable.choice)
ext.hpi.fevd = plot_fevd(position, ext.data.irf_fevd, parameters)
```

Under the extended model specification, the domestic monetary policy shock appears less important, whereas the home loan rate shock becomes much more important in explaining the variability in house price index. 

## Extension 2: Inverse Normal Gamma 2 Prior
```{r compute the irf and fevd results for data in extended model 2, eval=FALSE}
ext2.data.irf_fevd = irf_fevd(parameters, mydata, ext2.struct)
```

```{r save the irf and fevd results for data in extended model 2, eval=FALSE}
save(ext2.data.irf_fevd, file = "ext2.data.irf_fevd.RData")
```

```{r load the irf and fevd results for data in extended model 2, eval=TRUE}
load("ext2.data.irf_fevd.RData")
```

### Impulse Response Functions
```{r irf of cashrate extended model 2}
#| label: ext.cashrate.irf 2
#| fig-cap: "IRF plots for the AU Cash Rate Shock, Extended Model"
#| fig-cap-location: top
ext2.cashrate.irf = plot_irf(5, ext2.data.irf_fevd, parameters)
```
Some IRF results are different compared to the baseline results. The results also exhibit greater volatility which might be the result of the particular prior selection.

The effect of a positive domestic monetary policy shock on home loan rates seems to be more persistent as the effect erodes away at a much lower rate. This effect essentially mirrors the effect of a positive domestic monetary policy shock on itself.

The quantity of new homes built is the qualitatively the same as under the baseline specification, except for the absence of an initial boost in the number of houses built. Moreover, the quantity of new homes reverts back to its original levels by 14 quarters.

House price index results are also qualitatively the same as under the baseline specification. However, the magnitude of the reduction in house prices due to a positive monetary policy shock is higher than under the baseline specification but reverts back to a less lower level.

Moreover, a positive domestic monetary policy surprisingly seems to lower the money supply under this extension, which is not quite intuitive.

For the extended model, using the `plot_irf` function, we plot the IRFs to a +25 basis points increase in the US Federal Funds Rate.
```{r irf of usffr extended model 2}
#| label: ext.usffr.irf 2
#| fig-cap: "IRF plots for the US Federal Funds Rate Shock, Extended Model"
#| fig-cap-location: top
ext2.usffr.irf = plot_irf(1, ext2.data.irf_fevd, parameters)
```
These results qualitatively mirror the baseline results albeit exhibit greater volatility across periods.

### Forecast Error Variance Decomposition
```{r fevd for homeloan rates extended model 2, warning=FALSE}
variable.choice = "Owner-occupied home loan rates"
position = which(colnames(df) == variable.choice)
ext2.hlr.fevd = plot_fevd(position, ext2.data.irf_fevd, parameters)
```

The FEVD graph above shows that the domestic monetary policy shock carries the bulk of the explanatory power in the variability of home loan rates at the beginning but this power wanes over time. Compared to the baseline model, the shocks on real economy variables (GDP, consumption) become more  important over time.

```{r fevd for quantity of new homes extended model 2}
variable.choice = "Quantity of new homes (log)"
position = which(colnames(df) == variable.choice)
ext2.nh.fevd = plot_fevd(position, ext2.data.irf_fevd, parameters)
```

The FEVD graph above shows that while both domestic and foreign monetary policy have little explanatory power on the variability of the quantity of new homes built, the power increases over time. However, this increase in explanatory power is lower than in the baseline model. Moreover, shocks on the real economic indicators (GDP, consumption) and change in inflation are relatively more important in explaining the variability of the quantity of new homes.

```{r fevd for house price index extended model 2}
variable.choice = "House Price Index (log)"
position = which(colnames(df) == variable.choice)
ext2.hpi.fevd = plot_fevd(position, ext2.data.irf_fevd, parameters)
```
The FEVD for House Price Index shows that domestic and foreign policy shock becomes more important in explaining the variability over time as does aggregate consumption.

## Extension 3: Normal Gamma Prior with Stochastic Volatility
```{r compute the irf and fevd results for data in extended model 3, eval=FALSE}
ext3.data.irf_fevd = irf_fevd(parameters, mydata, ext3.struct)
```

```{r save the irf and fevd results for data in extended model 3, eval=FALSE}
save(ext3.data.irf_fevd, file = "ext3.data.irf_fevd.RData")
```

```{r load the irf and fevd results for data in extended model 3, eval=TRUE}
load("ext3.data.irf_fevd.RData")
```

### Impulse Response Functions
```{r irf of cashrate extended model 3}
#| label: ext.cashrate.irf 3
#| fig-cap: "IRF plots for the AU Cash Rate Shock, Extension 3"
#| fig-cap-location: top
ext3.cashrate.irf = plot_irf(5, ext3.data.irf_fevd, parameters)
```
The IRF results here are mostly similar to those obtained in extension 1. One key change that occurs is that the effect of domestic monetary policy shock on US FFR is now positive, much in line with expectations. 


For the extended model, using the `plot_irf` function, we plot the IRFs to a +25 basis points increase in the US Federal Funds Rate.
```{r irf of usffr extended model 3}
#| label: ext.usffr.irf 3
#| fig-cap: "IRF plots for the US Federal Funds Rate Shock, Extension 3"
#| fig-cap-location: top
ext3.usffr.irf = plot_irf(1, ext3.data.irf_fevd, parameters)
```
Again, the IRF results here are mostly similar to those obtained in extension 1. One key change that occurs is that the effect of US monetary policy shock on Australian cash rate is initially negative but bounces back to being positive within the span of 4 quarters. 

### Forecast Error Variance Decomposition
```{r fevd for homeloan rates extended model 3, warning=FALSE}
variable.choice = "Owner-occupied home loan rates"
position = which(colnames(df) == variable.choice)
ext3.hlr.fevd = plot_fevd(position, ext3.data.irf_fevd, parameters)
```

The FEVD graph above shows that the domestic monetary policy shock carries the bulk of the explanatory power in the variability of home loan rates, while foreign monetary policy shock is also important.

```{r fevd for quantity of new homes extended model 3}
variable.choice = "Quantity of new homes (log)"
position = which(colnames(df) == variable.choice)
ext3.nh.fevd = plot_fevd(position, ext3.data.irf_fevd, parameters)
```

The FEVD graph above shows that while both domestic and foreign monetary policy have little explanatory power on the variability of the quantity of new homes built, the power increases over time. However, this increase in explanatory power is lower than in the baseline model. House Price Index is much more important in explaining the variability of the quantity of new homes.

```{r fevd for house price index extended model 3}
variable.choice = "House Price Index (log)"
position = which(colnames(df) == variable.choice)
ext3.hpi.fevd = plot_fevd(position, ext3.data.irf_fevd, parameters)
```
The FEVD graph above shows that domestic and foreign monetary policy shocks have little explanatory power but become slightly more important over time in explaining the variability in housing prices. 

# Conclusion
In this paper, I investigate the effects of domestic and foreign monetary policy shocks on key indicators of the housing market i.e. the number of new houses built, home loan rates, and housing prices. I employ a structural VAR (SVAR) model, impose exclusion restrictions for identification, and use @waggoner2003a method of sampling for structural parameters. Further, I enrich my baseline model by incorporating hyperparameter estimations using two different set of priors, and also by enriching common stochastic volatility of structural errors. In general, I find that a positive 25 basis point increase in the Australian Cash Rate reduces both the number of new houses and housing prices over the long run, while a positive foreign (US) monetary policy shock reduces the number of new houses but increases housing prices.

# Appendix {#appendix}
```{r, file="L16 codes.R"}
```

## Baseline Model
### Forecast Error Variance Decomposition (FEVD)
```{r fevd for nominal gdp}
#| code-fold: true
#| code-summary: "Baseline Model: FEVD for nominal GDP"
variable.choice = "Nominal GDP (log)"
position = which(colnames(df) == variable.choice)
gdp.fevd = plot_fevd(position, data.irf_fevd, parameters)
```

```{r fevd for consumption}
variable.choice = "Aggregate Consumption (log)"
position = which(colnames(df) == variable.choice)
con.fevd = plot_fevd(position, data.irf_fevd, parameters)
```

```{r fevd for change in cpi}
variable.choice = "% Change in CPI, yoy"
position = which(colnames(df) == variable.choice)
cpi.fevd = plot_fevd(position, data.irf_fevd, parameters)
```

```{r fevd for usffr}
variable.choice = "US FFR Max Target Rate"
position = which(colnames(df) == variable.choice)
usffr.fevd = plot_fevd(position, data.irf_fevd, parameters)
```

```{r fevd for cashrate, warning=FALSE}
variable.choice = "RBA Target Cash Rate"
position = which(colnames(df) == variable.choice)
aucr.fevd = plot_fevd(position, data.irf_fevd, parameters)
```


```{r fevd for m1}
variable.choice = "M1 Money Supply (logs)"
position = which(colnames(df) == variable.choice)
m1.fevd = plot_fevd(position, data.irf_fevd, parameters)
```


```{r save fevd figures in a single plot, eval=FALSE}
fevd.plots = c("usffr.fevd", "gdp.fevd", "con.fevd", "cpi.fevd", "aucr.fevd", "m1.fevd", "hlr.fevd", "nh.fevd", "hpi.fevd")
  # Save each plot as an EPS file
  for (i in 1:length(fevd.plots)) {
    filename <- paste0(fevd.plots[i], ".pdf")
    pdf(file = filename, onefile = FALSE, height=9, width = 16)
    plot_fevd(i, data.irf_fevd, parameters)
    dev.off()
  }

```

### Extended Model
```{r fevd for nominal gdp, extended model}
variable.choice = "Nominal GDP (log)"
position = which(colnames(df) == variable.choice)
ext.gdp.fevd = plot_fevd(position, ext.data.irf_fevd, parameters)
```

```{r fevd for consumption, extended model}
variable.choice = "Aggregate Consumption (log)"
position = which(colnames(df) == variable.choice)
ext.con.fevd = plot_fevd(position, ext.data.irf_fevd, parameters)
```

```{r fevd for change in cpi, extended model}
variable.choice = "% Change in CPI, yoy"
position = which(colnames(df) == variable.choice)
ext.cpi.fevd = plot_fevd(position, ext.data.irf_fevd, parameters)
```

```{r fevd for usffr, extended model}
variable.choice = "US FFR Max Target Rate"
position = which(colnames(df) == variable.choice)
ext.usffr.fevd = plot_fevd(position, ext.data.irf_fevd, parameters)
```


```{r fevd for cashrate, extended model, warning=FALSE}
variable.choice = "RBA Target Cash Rate"
position = which(colnames(df) == variable.choice)
ext.aucr.fevd = plot_fevd(position, ext.data.irf_fevd, parameters)
```


```{r fevd for m1, extended model}
variable.choice = "M1 Money Supply (logs)"
position = which(colnames(df) == variable.choice)
ext.m1.fevd = plot_fevd(position, ext.data.irf_fevd, parameters)
```

```{r save fevd figures in a single plot for extended model, eval=FALSE}
fevd.plots = c("ext.usffr.fevd", "ext.gdp.fevd", "ext.con.fevd", "ext.cpi.fevd", "ext.aucr.fevd", "ext.m1.fevd", "ext.hlr.fevd", "ext.nh.fevd", "ext.hpi.fevd")
  # Save each plot as an EPS file
  for (i in 1:length(fevd.plots)) {
    filename <- paste0(fevd.plots[i], ".pdf")
    pdf(file = filename, onefile = FALSE, height=9, width = 16)
    plot_fevd(i, ext.data.irf_fevd, parameters)
    dev.off()
  }

```

### Extension 2
```{r fevd for nominal gdp extended model 2}
variable.choice = "Nominal GDP (log)"
position = which(colnames(df) == variable.choice)
ext2.gdp.fevd = plot_fevd(position, ext2.data.irf_fevd, parameters)
```

```{r fevd for consumption extended model 2}
variable.choice = "Aggregate Consumption (log)"
position = which(colnames(df) == variable.choice)
ext2.con.fevd = plot_fevd(position, ext2.data.irf_fevd, parameters)
```

```{r fevd for change in cpi extended model 2}
variable.choice = "% Change in CPI, yoy"
position = which(colnames(df) == variable.choice)
ext2.cpi.fevd = plot_fevd(position, ext2.data.irf_fevd, parameters)
```

```{r fevd for usffr extended model 2}
variable.choice = "US FFR Max Target Rate"
position = which(colnames(df) == variable.choice)
ext2.usffr.fevd = plot_fevd(position, ext2.data.irf_fevd, parameters)
```


```{r fevd for cashrate extended model 2, warning=FALSE}
variable.choice = "RBA Target Cash Rate"
position = which(colnames(df) == variable.choice)
ext2.aucr.fevd = plot_fevd(position, ext2.data.irf_fevd, parameters)
```


```{r fevd for m1 extended model 2}
variable.choice = "M1 Money Supply (logs)"
position = which(colnames(df) == variable.choice)
ext2.m1.fevd = plot_fevd(position, ext2.data.irf_fevd, parameters)
```

```{r save fevd figures in a single plot for extended model 2, eval=FALSE}
fevd2.plots = c("ext2.usffr.fevd", "ext2.gdp.fevd", "ext2.con.fevd", "ext2.cpi.fevd", "ext2.aucr.fevd", "ext2.m1.fevd", "ext2.hlr.fevd", "ext2.nh.fevd", "ext2.hpi.fevd")
  # Save each plot as an EPS file
  for (i in 1:length(fevd2.plots)) {
    filename <- paste0(fevd2.plots[i], ".pdf")
    pdf(file = filename, onefile = FALSE, height=9, width = 16)
    plot_fevd(i, ext2.data.irf_fevd, parameters)
    dev.off()
  }

```

### Extension 3
```{r fevd for nominal gdp extended model 3}
variable.choice = "Nominal GDP (log)"
position = which(colnames(df) == variable.choice)
ext3.gdp.fevd = plot_fevd(position, ext3.data.irf_fevd, parameters)
```

```{r fevd for consumption extended model 3}
variable.choice = "Aggregate Consumption (log)"
position = which(colnames(df) == variable.choice)
ext3.con.fevd = plot_fevd(position, ext3.data.irf_fevd, parameters)
```

```{r fevd for change in cpi extended model 3}
variable.choice = "% Change in CPI, yoy"
position = which(colnames(df) == variable.choice)
ext3.cpi.fevd = plot_fevd(position, ext3.data.irf_fevd, parameters)
```

```{r fevd for usffr extended model 3}
variable.choice = "US FFR Max Target Rate"
position = which(colnames(df) == variable.choice)
ext3.usffr.fevd = plot_fevd(position, ext3.data.irf_fevd, parameters)
```


```{r fevd for cashrate extended model 3, warning=FALSE}
variable.choice = "RBA Target Cash Rate"
position = which(colnames(df) == variable.choice)
ext3.aucr.fevd = plot_fevd(position, ext3.data.irf_fevd, parameters)
```


```{r fevd for m1 extended model 3}
variable.choice = "M1 Money Supply (logs)"
position = which(colnames(df) == variable.choice)
ext3.m1.fevd = plot_fevd(position, ext3.data.irf_fevd, parameters)
```

```{r save fevd figures in a single plot for extended model 3, eval=FALSE}
fevd3.plots = c("ext3.usffr.fevd", "ext3.gdp.fevd", "ext3.con.fevd", "ext3.cpi.fevd", "ext3.aucr.fevd", "ext3.m1.fevd", "ext3.hlr.fevd", "ext3.nh.fevd", "ext3.hpi.fevd")
  # Save each plot as an EPS file
  for (i in 1:length(fevd3.plots)) {
    filename <- paste0(fevd3.plots[i], ".pdf")
    pdf(file = filename, onefile = FALSE, height=9, width = 16)
    plot_fevd(i, ext3.data.irf_fevd, parameters)
    dev.off()
  }

```

## Code Snippets
```{r}
#| echo: true
#| code-fold: true
#| code-summary: "Code: Orthogonal Complement Matrix Function"
orthogonal.complement.matrix.TW = function(x){
  # x is a mxn matrix and m>n
  # the function returns a mx(m-n) matrix, out, that is an orthogonal complement of x, i.e.:
  # t(x)%*%out = 0 and det(cbind(x,out))!=0
  if( dim(x)[1] == 1 & dim(x)[2] == 2){
    x = t(x)
  }
  # x <- ifelse(dim(x)[1] == 1 && dim(x)[2] == 2, t(x), x)
  N     = dim(x)
  tmp   = qr.Q(qr(x, tol = 1e-10),complete=TRUE)
  out   = as.matrix(tmp[,(N[2]+1):N[1]])
  return(out)
}
```

```{r}
#| echo: true
#| code-fold: true
#| code-summary: "Code: Sample from Conditional Generalised Normal Dist Function"
r.conditional.generalized.normal = function(S.inv, nu, Vn, n, B0){
  # A function to sample a random draw from a conditional generalized normal distribution
  # of the unrestricted elements of the n-th row of matrix B0 
  # given the parameters from the remaining rows of B0
  # Depends on package mvtnorm
  # use: library(rmvtnorm)
  
  rn            = nrow(Vn)
  Un            = chol(nu*solve(Vn%*%S.inv%*%t(Vn)))
  w             = t(orthogonal.complement.matrix.TW(t(B0[-n,])))
  w1            = w %*% t(Vn) %*% t(Un) / sqrt(as.numeric(w %*% t(Vn) %*% t(Un) %*% Un %*% Vn %*% t(w)))
  if (rn>1){
    Wn          = cbind(t(w1),orthogonal.complement.matrix.TW(t(w1)))
  } else {
    Wn          = w1
  }
  alpha         = rep(NA,rn)
  u             = rmvnorm(1,rep(0,nu+1),(1/nu)*diag(nu+1))
  alpha[1]      = sqrt(as.numeric(u%*%t(u)))
  if (runif(1)<0.5){
    alpha[1]    = -alpha[1]
  }
  if (rn>1){
    alpha[2:rn] = rmvnorm(1,rep(0,nrow(Vn)-1),(1/nu)*diag(rn-1))
  }
  bn            = alpha %*% Wn %*% Un
  B0n           = bn %*% Vn
  
  output        = list(bn=bn, B0n=B0n)
  return(output)
}
```

```{r}
#| echo: true
#| code-fold: true
#| code-summary: "Code: Sample B0 Function"
rgn             = function(n,S.inv,nu,V,B0.initial){
  # This function simulates draws for the unrestricted elements 
  # of the conteporaneous relationships matrix of an SVAR model
  # from a generalized-normal distribution according to algorithm 
  # by Waggoner & Zha (2003, JEDC)
  # n     - a positive integer, the number of draws to be sampled
  # S     - an NxN positive definite matrix, a parameter of the generalized-normal distribution
  # nu    - a positive scalar, degrees of freedom parameter
  # V     - an N-element list, with fixed matrices
  # B0.initial - an NxN matrix, of initial values of the parameters
  
  N             = nrow(B0.initial)
  no.draws      = n
  
  B0            = array(NA, c(N,N,no.draws))
  B0.aux        = B0.initial
  
  for (i in 1:no.draws){
    for (n in 1:N){
      rn            = nrow(V[[n]])
      Un            = chol(nu*solve(V[[n]]%*%S.inv%*%t(V[[n]])))
      w             = t(orthogonal.complement.matrix.TW(t(B0.aux[-n,])))
      w1            = w %*% t(V[[n]]) %*% t(Un) / sqrt(as.numeric(w %*% t(V[[n]]) %*% t(Un) %*% Un %*% V[[n]] %*% t(w)))
      if (rn>1){
        Wn          = cbind(t(w1),orthogonal.complement.matrix.TW(t(w1)))
      } else {
        Wn          = w1
      }
      alpha         = rep(NA,rn)
      u             = rmvnorm(1,rep(0,nu+1),(1/nu)*diag(nu+1))
      alpha[1]      = sqrt(as.numeric(u%*%t(u)))
      if (runif(1)<0.5){
        alpha[1]    = -alpha[1]
      }
      if (rn>1){
        alpha[2:rn] = rmvnorm(1,rep(0,nrow(V[[n]])-1),(1/nu)*diag(rn-1))
      }
      bn            = alpha %*% Wn %*% Un
      B0.aux[n,]    = bn %*% V[[n]]
    }
    B0[,,i]         = B0.aux
  }
  
  return(B0)
}
```


```{r}
#| echo: true
#| code-fold: true
#| code-summary: "Code: Normalise B0 Function"
normalization.wz2003  = function(B0,B0.hat.inv, Sigma.inv, diag.signs){
  # This function normalizes a matrix of contemporaneous effects
  # according to the algorithm by Waggoner & Zha (2003, JOE)
  # B0        - an NxN matrix, to be normalized
  # B0.hat    - an NxN matrix, a normalized matrix
  
  N                 = nrow(B0)
  K                 = 2^N
  distance          = rep(NA,K)
  for (k in 1:K){
    B0.tmp.inv      = solve(diag(diag.signs[k,]) %*% B0)
    distance[k]     = sum(
      unlist(
        lapply(1:N,
               function(n){
                 t(B0.tmp.inv - B0.hat.inv)[n,] %*%Sigma.inv %*% t(B0.tmp.inv - B0.hat.inv)[n,]
               }
        )))
  }
  B0.out            = diag(diag.signs[which.min(distance),]) %*% B0
  
  return(B0.out)
}
```

```{r}
#| echo: true
#| code-fold: true
#| code-summary: "Code: Normalise B0 Parallelised Function"
normalize.Gibbs.output.parallel          = function(B0.posterior,B0.hat){
  # This function normalizes the Gibbs sampler output from function rgn
  # using function normalization.wz2003 
  # B0.posterior  - a list, output from function rgn
  # B0.hat        - an NxN matrix, a normalized matrix
  
  N                 = nrow(B0.hat)
  K                 = 2^N
  
  B0.hat.inv        = solve(B0.hat)
  Sigma.inv         = t(B0.hat)%*%B0.hat
  
  diag.signs        = matrix(NA,2^N,N)
  for (n in 1:N){
    diag.signs[,n]  = kronecker(c(-1,1),rep(1,2^(n-1)))
  }
  
  B0.posterior.n    = mclapply(1:dim(B0.posterior)[3],function(i){
    normalization.wz2003(B0=B0.posterior[,,i],B0.hat.inv, Sigma.inv, diag.signs)
  },mc.cores=6
  )
  B0.posterior.n  = simplify2array(B0.posterior.n)
  
  return(B0.posterior.n)
}
```

```{r}
normalize.Gibbs.output          = function(B0.posterior,B0.hat){
  # This function normalizes the Gibbs sampler output from function rgn
  # using function normalization.wz2003 
  # B0.posterior  - a list, output from function rgn
  # B0.hat        - an NxN matrix, a normalized matrix
  
  N                 = nrow(B0.hat)
  K                 = 2^N
  
  B0.hat.inv        = solve(B0.hat)
  Sigma.inv         = solve(B0.hat.inv %*% t(B0.hat.inv))
  
  diag.signs        = matrix(NA,2^N,N)
  for (n in 1:N){
    diag.signs[,n]  = kronecker(c(-1,1),rep(1,2^(n-1)))
  }
  
  for (i in 1:dim(B0.posterior)[3]){
    if (i%%100==0){ cat(i," ")}
    norm.post                   = normalization.wz2003(B0=B0.posterior[,,i],B0.hat.inv, Sigma.inv, diag.signs)
    B0.posterior[,,i]           = norm.post
  }
  return(B0.posterior)
}
```

```{r}
#| echo: true
#| code-fold: true
#| code-summary: "Code: Sample Bplus Function"
rnorm.ngn       = function(B0.posterior,B,Omega){
  # This function simulates draws for the multivariate normal distribution
  # of the autoregressive slope matrix of an SVAR model
  # from a normal-generalized-normal distribution according to algorithm 
  # by Waggoner & Zha (2003, JEDC)
  # B0.posterior  - a list, output from function rgn
  # B             - an NxK matrix, a parameter determining the mean of the multivariate conditionally normal distribution given B0
  # Omega         - a KxK positive definite matrix, a covariance matrix of the multivariate normal distribution
  
  N             = nrow(B)
  K             = ncol(B)
  no.draws      = dim(B0.posterior)[3]
  L             = t(chol(Omega))
  
  Bp.posterior  = lapply(1:no.draws,function(i){
    Bp          = matrix(NA, N, K)
    for (n in 1:N){
      Bp[n,]    = as.vector(t(B0.posterior[n,,i] %*% B) + L%*%rnorm(K))
    }
    return(Bp)
  })
  Bp.posterior  = simplify2array(Bp.posterior)
  return(Bp.posterior)
}

```


# References {.unnumbered}



